---
title: "01 - Data exploration
subtitle: "Airline Customer Satisfaction -- Bayesian Learning & Monte Carlo Simulations, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(caret)
library(RColorBrewer)

rm(list = ls())
set.seed(2026)
```

## 0. Dataset import

```{r}
# dataset loading

raw <- read.csv("../data/Airline_customer_satisfaction.csv", stringsAsFactors = FALSE)

# manually set "satisfied" class encoding to 1 (later use)
raw <- raw %>%
  mutate(
    satisfaction   = factor(satisfaction, levels = c("dissatisfied", "satisfied")),
    # even if for this variables their factor won't be relevant during prediction,  is still useful for data exploration
    Class          = factor(Class, levels = c("Eco", "Eco Plus", "Business")),
    Customer.Type  = factor(Customer.Type),
    Type.of.Travel = factor(Type.of.Travel)
  )

```

In order to make the columns easier to manager, we change the column's names, switching to `camel_case` notation.

```{r}
new_names <- c(
  "satisfaction", "customer_type", "age", "travel_type", "cabin_class",
  "flight_distance", "seat_comfort", "time_convenient", "food_drink",
  "gate_location", "wifi", "entertainment", "online_support", "ease_booking",
  "onboard_service", "legroom", "baggage", "checkin", "cleanliness",
  "online_boarding", "dep_delay", "arr_delay"
)

names(raw) <- new_names
dim(raw)
```

In order to avoid later problems, we immediately check for missing values:

```{r}
# find and locate missing values
na_counts <- colSums(is.na(raw))
na_counts[na_counts > 0]
```

The only column missing values is (`r na_counts["arr_delay"]` rows, `r round(100*na_counts["arr_delay"]/nrow(raw), 2)`% of the data). Given the small percentage of missing values, we decided to drop those records sine probably it won't impact the outcome.

```{r}
nrow(raw)
raw <- raw %>% filter(!is.na(arr_delay))
raw$row_id <- seq_len(nrow(raw))
nrow(raw)
```

## 1. Dataset exploration

Explore general dataset properties.

```{r}
# general info
str(raw)

# check for duplicated rows
cat("Duplicated rows:", sum(duplicated(raw)), "\n")
```

Now we check variable domain, especially whether all service-quality items lay within the range 0-5.

```{r}
col_filter = c("age", "flight_distance", "seat_comfort", "time_convenient", "food_drink", "gate_location", "wifi", "entertainment", "online_support", "ease_booking", "onboard_service", "legroom", "baggage", "checkin", "cleanliness", "online_boarding", "dep_delay")

sapply(raw[col_filter], range)
```

All quality service variables range over 0-5. The remaining numeric variables also look sane: `age` in [`r min(raw$age)`, `r max(raw$age)`], `flight_distance` in [`r min(raw$flight_distance)`, `r max(raw$flight_distance)`] miles, except for `arr_delay` due the presence f NaN.

### 1.1 Target variable class balance

```{r}
# check and lot target class balance

pal <- c(dissatisfied = "#eb6834", satisfied = "#2a78d0")

bal <- raw %>%
  count(satisfaction) %>%
  mutate(share = n / sum(n),
         satisfaction = factor(satisfaction, levels = c("dissatisfied", "satisfied")))

ggplot(bal, aes(satisfaction, n, fill = satisfaction)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = sprintf("%d\n(%.1f%%)", n, 100 * share)),
            vjust = -0.3, size = 4, colour = "#52514e", lineheight = 0.95) +
  scale_fill_manual(values = pal) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(title = "Class balance of the target variable",
       x = NULL, y = "customers") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank())
```

The response is close to balanced (`r round(100*mean(raw$satisfaction == "satisfied"),1)`% satisfied), therefore we won't need something more complex than a Bernoulli+logistic regression model.

### 1.2 Boxplots

```{r}
# we modify categorical variables in order to show them in the boxplots and correlation matrix

num.raw = raw
cat_vars = names(num.raw)[!sapply(num.raw, is.numeric)]

mappings <- lapply(num.raw[cat_vars], levels)
num.raw[cat_vars] <- lapply(num.raw[cat_vars], as.integer)
```

```{r fig.width=16, fig.height=14}
num.raw %>%
  select(where(is.numeric)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(show.legend = FALSE) +
  facet_wrap(~ variable, scales = "free") +
  scale_fill_viridis_d() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

### 1.2.1 Quality of Service indicators by satisfaction

```{r fig.width=14, fig.height=14}
qs_vars <- c("seat_comfort", "time_convenient", "food_drink", "gate_location", "wifi", "entertainment", "online_support", "ease_booking", "onboard_service", "legroom", "baggage", "checkin", "cleanliness", "online_boarding")

raw %>%
  select(satisfaction, all_of(qs_vars)) %>%
  pivot_longer(-satisfaction, names_to = "item", values_to = "rating") %>%
  ggplot(aes(x = satisfaction, y = rating, fill = satisfaction)) +
  geom_boxplot(outlier.size = 0.4, alpha = 0.85) +
  facet_wrap(~ item, ncol = 4, scales = "free_y") +
  scale_fill_manual(values = c("dissatisfied" = "#D85A30", "satisfied" = "#3B6D11")) +
  labs(title = "Service-quality ratings by satisfaction outcome",
       x = NULL, y = "Rating (0-5)") +
  theme_minimal(base_size = 9) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 0),
        strip.text = element_text(face = "bold", size = 7.5))
```

The customer satisfaction per item follows the same pattern expressed by the correlation between the QS variable and the respective correlation (see below). Generally speaking we can see that the higher the customer satisfaction the higher the QS was.

### 1.3 Full correlation heatmap

```{r fig.width=16, fig.height=16}

corr_mat = cor(num.raw[, !names(num.raw) == "row_id"])

corrplot(corr_mat, method = "color", type = "upper", order = "hclust",
         col = colorRampPalette(rev(brewer.pal(9, "RdBu")))(200),
         tl.col = "black", tl.srt = 45, tl.cex = 0.75,
         addCoef.col = "black", number.cex = 0.55,
         diag = FALSE,
         title = "Correlation matrix of all original and augmented covariates",
         mar = c(0, 0, 2, 0))
```

### 1.3.1 Quality service correlation

```{r fig.width=14, fig.height=14}
# focus on the correlation between Quality of Service variables

# re-defining them for clarity
qs_vars <- c("seat_comfort", "time_convenient", "food_drink", "gate_location", "wifi", "entertainment", "online_support", "ease_booking", "onboard_service", "legroom", "baggage", "checkin", "cleanliness", "online_boarding")


qs_corr <- cor(num.raw[, qs_vars], use = "pairwise.complete.obs")

corrplot(qs_corr, method = "color", type = "upper", order = "hclust",
         col = colorRampPalette(rev(brewer.pal(9, "RdBu")))(200),
         tl.col = "black", tl.srt = 45, tl.cex = 0.75,
         addCoef.col = "black", number.cex = 0.55,
         diag = FALSE,
         title = "Correlation matrix of the 14 service-quality variables",
         mar = c(0, 0, 2, 0))
```

from the quality of service correlation matrix we can distinguish 3 clusters of variables.

```{r}
cluster_range <- function(items) {
  sub <- qs_corr[items, items]
  vals <- sub[upper.tri(sub)]
  sprintf("%.2f-%.2f", min(vals), max(vals))
}


cluster_booking  <- c("wifi", "entertainment", "online_support", "ease_booking")

cluster_onboard  <- c("onboard_service", "legroom", "baggage")

cluster_catering <- c("seat_comfort", "food_drink", "gate_location")

r_booking  <- cluster_range(cluster_booking)
r_onboard  <- cluster_range(cluster_onboard)
r_catering <- cluster_range(cluster_catering)

r_booking
r_onboard
r_catering
```

- **Booking / connectivity**: `Inflight wifi service`, `Online support`, `Ease of Online booking`, `Online boarding` move together (pairwise correlations `r r_booking`).
- **On-board comfort**: `On-board service`, `Leg room service`, `Baggage handling`, `Cleanliness` form a second cluster (`r r_onboard`).
- **Seat / catering / ground experience**: `Seat comfort`, `Food and drink`, `Departure/Arrival time convenient`, `Gate location` form a third, slightly tighter cluster (`r r_catering`).

### 1.4 Structural zeros in quality of service variables

Several quality of service show a suspiciously large share of exact `0` ratings, which plausibly encode "not applicable" (e.g. no on-board wifi on some flights) rather than a genuine bottom-of-scale rating.

```{r structural-zeros}
zero_share <- sort(sapply(raw[, qs_vars], function(x) mean(x == 0)),
                    decreasing = TRUE)

knitr::kable(
  data.frame(item = names(zero_share), zero_share = round(zero_share, 3)),
  row.names = FALSE, col.names = c("Service item", "Share of rows = 0"),
  caption = "Share of rows recording a 0 rating, by service item"
)
```

We flag, but do not model, the items with a non-trivial zero share (above 2%) as a single per-row indicator, to be used later only as a robustness check

```{r}
zero_candidates <- names(zero_share)[zero_share > 0.02]

raw <- raw %>%
  mutate(structural_zero_flag = if_any(all_of(zero_candidates), ~ .x == 0))

n_flagged <- sum(raw$structural_zero_flag)
n_flagged
```

## 2. Data augmentation

Before any exploration we show some changes we've done to the dataset, avoiding to show irrelevant information.

### 2.1 Decorrelating Departure and Arrival delay

```{r delay-collinearity}
delay_cor <- cor(raw$dep_delay, raw$arr_delay, use = "complete.obs")
n_missing_arr <- sum(is.na(raw$arr_delay))
```

`Departure Delay` and `Arrival Delay` correlate at `r sprintf("%.3f", delay_cor)`: using both raw would be close to duplicating one variable.

We tought about two solutions:

- taking the difference between them and accounting for the "mid-fly delay"

- enriching the previous concept and regress `Arrival Delay` on `Departure Delay` and keep the **residual** as a second covariate -- by construction uncorrelated with `Departure Delay`, and interpretable as "minutes gained or lost in the air relative to what the departure delay alone would predict":

```{r delay-residual}
# exploring the 

delay_fit <- lm(arr_delay ~ dep_delay, data = raw)
raw$arr_delay_res <- residuals(delay_fit)

check_cor <- cor(raw$dep_delay, raw$arr_delay, use = "complete.obs")
```

The fitted relationship is Arrival $\approx$ `r sprintf("%.2f", coef(delay_fit)[1])` $+$ `r sprintf("%.3f", coef(delay_fit)[2])` $\times$ Departure (R$^2$ = `r sprintf("%.3f", summary(delay_fit)$r.squared)`).

Given that the slope is almost one and the intercept is just `r`coef(delay_fit)[1]\` minutes, we opted for the simple difference as solution, given the simplicity and higher interpretability.

```{r}
raw$midfly_delay <- raw$arr_delay - raw$dep_delay
```

### 2.1 Adjust for skewnees

From the boxplots we can see that covariates `flight distance`, `departure delay' and`midfly delay\` are skew. So we apply a log-transformation and we check the results to verify if the transformation is worthy or not.

```{r skew-check}
skew <- function(x) { x <- x[!is.na(x)]; m <- mean(x); s <- sd(x); mean(((x - m) / s)^3) }

signedLogArr <- sign(raw$midfly_delay) * log1p(abs(raw$midfly_delay))

skew_tbl <- data.frame(
  variable = c("Flight Distance", "log1p(Flight Distance)",
               "Departure Delay", "log1p(Departure Delay)",
               "Midflight Delay", "signed-log(Midflight delay)"
               ),
  skewness = round(c(
    skew(raw$flight_distance), skew(log1p(raw$flight_distance)), 
    skew(raw$dep_delay), skew(log1p(raw$dep_delay)),
    skew(raw$midfly_delay), skew(signedLogArr)
    ), 3)
)
knitr::kable(skew_tbl)
```

Flight Distance is only mildly skewed (`r`skew(raw$flight_distance)`) and a log transform overcorrects it into a left skew (`r `skew(log1p(raw$flight_distance`) -- so we keep it on the raw scale. The two delay variables are genuinely skewed (`r `skew(raw$dep_delay)` and `r`skew(raw\$midfly_delay)\`) and both transforms clearly help: log(1+x) for Departure Delay (non-negative), and a signed log, $\text{sign}(x)\log(1+|x|)$, for the delay residual, which can be negative.

```{r}
raw$log_mid_delay <- signedLogArr
raw$log_dep_delay <- log1p(raw$dep_delay)
```

### 2.3 Final correlation matrix

```{r fig.width=16, fig.height=16}

new_corr_mat = cor(num.raw[, !names(num.raw) == "row_id"])
# arr_delay_res is excluded due problems during the plotting

corrplot(new_corr_mat, method = "color", type = "upper", order = "hclust",
         col = colorRampPalette(rev(brewer.pal(9, "RdBu")))(200),
         tl.col = "black", tl.srt = 45, tl.cex = 0.75,
         addCoef.col = "black", number.cex = 0.55,
         diag = FALSE,
         title = "Correlation matrix of all original and augmented covariates",
         mar = c(0, 0, 2, 0))
```

## 3. Data stratification, split and storage

In the remaining part we stratify the dataset for the model, we split it and we store it for later use.

### 3.1 Data stratification and split

The brief describes a curated 1,000-row file; we have the full `r format(n_total, big.mark = ",")`-row source. We draw a stratified training subsample of $n \approx 1{,}200$, and we stratify it to reflect categorical variable classes proportions.

```{r fig.width=14, fig.height=14}

cat_vars <- c("satisfaction", "customer_type", "travel_type", "cabin_class")

lapply(raw[cat_vars], function(x) round(prop.table(table(x)), 3))
```

Basically all categorical variables need to be stratified.

```{r}
set.seed(123)

strata <- interaction(raw$satisfaction, raw$cabin_class, raw$customer_type, raw$travel_type, drop = TRUE)
train_idx <- createDataPartition(y = strata, p = 1200 / nrow(raw), list = FALSE)
  
train_df <- raw[train_idx, ]
test_df  <- raw[-train_idx, ]
```

Training subsample: `r nrow(train_df)` rows. Held-out test set: `r format(nrow(test_df), big.mark = ",")` rows.

```{r stratification-check}
compare_tbl <- function(var) {
  full  <- prop.table(table(raw[[var]]))
  train <- prop.table(table(train_df[[var]]))
  round(cbind(full = full, train = train), 3)
}

cat("Satisfaction proportions (full data vs. training subsample):\n")
print(compare_tbl("satisfaction"))

cat("\nClass proportions (full data vs. training subsample):\n")
print(compare_tbl("cabin_class"))

cat("\nCustomer Type proportions (full data vs. training subsample):\n")
print(compare_tbl("customer_type"))
```

### 3.2 Dataset storage

```{r save-outputs}
saveRDS(raw, "../data/airline_full_clean.rds")
write.csv(train_df, "../data/airline_train_n1200.csv", row.names = FALSE)
write.csv(test_df,  "../data/airline_test_holdout.csv", row.names = FALSE)
```
