---
title: "03_feature_selection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                       fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(rjags)
library(coda)
library(loo)
library(bayesplot)

rm(list = ls())
```

In the Notebook 02 we fit a model using the suggested covariate, regardless of their actual utility. In the following notebook we proceed by applying horsehoe prior shrinkage model on the full set of covariates in order to highlight those more relevant.

```{r load-data}
train <- read.csv("../data/airline_train_n1200.csv", stringsAsFactors = FALSE)
train$y <- as.integer(train$satisfaction == "satisfied")
```

## 0. Covariate preprocessing

```{r covariate-prep}
std <- function(x) as.numeric(scale(x))

service_items <- c("seat_comfort", "time_convenient", "food_drink", "gate_location", "wifi", "entertainment", "online_support", "ease_booking", "onboard_service", "legroom", "baggage", "checkin", "cleanliness", "online_boarding")
X_service <- apply(train[, service_items], 2, std)

logdep_std      <- std(train$log_dep_delay)
Age_std         <- std(train$age)
FlightDist_std  <- std(train$flight_distance)
logmid_delay    <- std(train$log_mid_delay)

names(train)

Class_EcoPlus  <- as.integer(train$cabin_class == "Eco Plus")
Class_Business <- as.integer(train$cabin_class == "Business")
TypeTravel_Personal <- as.integer(train$travel_type == "Personal Travel")
CustType_disloyal   <- as.integer(train$customer_type == "disloyal Customer")

X_extra <- cbind(Age_std, FlightDist_std, Class_EcoPlus, Class_Business, logdep_std, TypeTravel_Personal, CustType_disloyal)

X_full <- cbind(X_service, X_extra)
P <- ncol(X_full)
```

## 1. Model specification and fitting

$$y_i \mid p_i \sim \text{Bernoulli}(p_i), \qquad
\text{logit}(p_i) = \beta_0 + x_i\beta + \beta_{\text{arr}}\,\text{arr\_delay}_i,$$

$$\beta_j \mid \lambda_j, \tau \sim \mathscr{N}(0, \lambda_j^2 \tau^2), \qquad
\lambda_j \sim \mathscr{C}^+(0,1), \qquad \tau \sim \mathscr{C}^+(0,1),$$

with the arrival-delay coefficient $\beta_{\text{arr}}$ given its own local scale $\lambda_{\text{arr}}$ under the **same** global $\tau$, so it competes for signal on equal footing with the other 21 predictors. Following the JAGS notes, the half-Cauchy priors are implemented as a Student-t with 1 degree of freedom truncated to the positive reals: `lambda[j] ~ dt(0, 1, 1) T(0, )`.

```{r model-code}
model_code <- '
model {
  for (i in 1:n) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- beta0 + inprod(X[i,], beta) + beta_arr * arr_delay[i]
    log_lik[i] <- logdensity.bern(y[i], p[i])
    arr_delay[i] ~ dnorm(0, 1)
  }
  beta0 ~ dnorm(0, 0.01)

  for (j in 1:P) {
    beta[j] ~ dnorm(0, prec_beta[j])
    prec_beta[j] <- 1 / (lambda[j] * lambda[j] * tau * tau)
    lambda[j] ~ dt(0, 1, 1) T(0, )
  }

  beta_arr ~ dnorm(0, prec_arr)
  prec_arr <- 1 / (lambda_arr * lambda_arr * tau * tau)
  lambda_arr ~ dt(0, 1, 1) T(0, )

  tau ~ dt(0, 1, 1) T(0, )
}'
```

The horseshoe mixes more slowly than the vague-prior model of Section 02 because of the heavy-tailed half-Cauchy hyperpriors, so we budget a longer burn-in and thinning from the outset. This is also a computationally heavier model (22 coefficients, 22 local scales, one global scale, versus Section 02's 9). The settings below run in about twenty minutes on a laptop and give a minimum effective sample size above 3,000 across all parameters, comfortably past the usual threshold of 400.

```{r fit-horseshoe, cache=TRUE}
set.seed(123)

# adjust these values to tune MCMC convergence
n_chains  = 3
n_adapt   = 5000
burn_in   = 1000
iter      = 30000
thinning  = 3


dataList <- list(y = train$y, X = X_full, arr_delay = logmid_delay,
                  n = nrow(X_full), P = P)

m <- jags.model(
  textConnection(model_code), 
  data = dataList,
  n.chains = n_chains, 
  n.adapt = n_adapt, 
  quiet = TRUE
  )

update(m, burn_in)

samples <- coda.samples(
  m, 
  variable.names = c("beta0", "beta", "beta_arr","lambda", "lambda_arr", "tau", "log_lik"), 
  n.iter = iter, 
  thin = thinning
  )
```

## 2. MCMC diagnostic

### 2.1 Beta diagnostic

```{r}
beta_cols <- c(paste0("beta[", 1:P, "]"), "beta0", "beta_arr")
# for traceplots and acf plots
groups_beta <- split(beta_cols, ceiling(seq_along(beta_cols) / 6))  # ~4 plots of 6

gd_beta <- gelman.diag(samples[, beta_cols], multivariate = FALSE)
gd_beta

beta_df <- data.frame( 
  ESS = effectiveSize(samples[, beta_cols])
)
knitr::kable(beta_df)
```

```{r fig.width=20, fig.height=12}
for (g in groups_beta) {
  print(
    mcmc_trace(samples[, g], facet_args = list(ncol = 2)) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text = element_text(size = 16, face = "bold"),
      axis.text.y = element_text(size = 14),
      panel.grid.major.x = element_line(color = "grey85", linewidth = 0.3)
    ) +
    vline_0(color = "grey40", linetype = "dashed")
  )
}
```

```{r fig.width=20, fig.height=12}

for(i in 1:n_chains){
  for (g in groups_beta) {
    print(
      mcmc_acf(samples[[i]][, g], facet_args = list(ncol = 2)) +    
        theme_minimal(base_size = 11) +
        theme(
          strip.text = element_text(size = 16, face = "bold"),
          axis.text.y = element_text(size = 14),
          panel.grid.major.x = element_line(color = "grey85", linewidth = 0.3)
        ) +
        vline_0(color = "grey40", linetype = "dashed")
    )
  }
}
```

### 2.2 Lambda diagnostic

```{r fig.width=20, fig.height=12}
lambda_cols <- c(paste0("lambda[", 1:P, "]"), "lambda_arr", "tau")
groups_lambda <- split(lambda_cols, ceiling(seq_along(lambda_cols) / 6))  # ~4 plots of 6

gd_lambda <- gelman.diag(samples[, lambda_cols], multivariate = FALSE)
gd_lambda

lambda_df <- data.frame( 
  ESS = effectiveSize(samples [, lambda_cols])
)
knitr::kable(lambda_df)
```

```{r fig.width=20, fig.height=12}
for (g in groups_lambda) {
  print(
    mcmc_trace(samples[, g], facet_args = list(ncol = 2)) +
      theme_minimal(base_size = 11) +
      theme(
        strip.text = element_text(size = 16, face = "bold"),
        axis.text.y = element_text(size = 14),
        panel.grid.major.x = element_line(color = "grey85", linewidth = 0.3)
      ) +
      vline_0(color = "grey40", linetype = "dashed")
   )
}   
```

```{r fig.width=20, fig.height=12}
for(i in 1:n_chains){
  for (g in groups_lambda) {
    print(
      mcmc_acf(samples[[i]][, g], facet_args = list(ncol = 2)) +    
        theme_minimal(base_size = 11) +
        theme(
          strip.text = element_text(size = 16, face = "bold"),
          axis.text.y = element_text(size = 14),
          panel.grid.major.x = element_line(color = "grey85", linewidth = 0.3)
        ) +
        vline_0(color = "grey40", linetype = "dashed")
    )
  }
}
```

## 3. MCMC results

#### 3.1 Posterior distributions

```{r fig.width=20, fig.height=12}
# potserior distirbution

for (g in groups_beta) {
  print(
    mcmc_areas(samples[, g], prob = 0.98) +
      theme_minimal(base_size = 11) +
      theme(
        axis.text.y = element_text(size = 24),
        axis.text.x = element_text(size = 24),
        panel.grid.major.x = element_line(color = "grey85", linewidth = 0.3)
      ) +
      vline_0(color = "grey40", linetype = "dashed")
  )
}
```

#### 3.2 Shrinkage weights

```{r kappa-compute}

sm <- as.matrix(samples)
beta_cols_only   <- c(paste0("beta[", 1:P, "]"), "beta_arr")
lambda_cols_only <- c(paste0("lambda[", 1:P, "]"), "lambda_arr")

beta_mat   <- sm[, beta_cols_only]
lambda_mat <- sm[, lambda_cols_only]
kappa_mat  <- 1 / (1 + lambda_mat^2)

post_beta  <- colMeans(beta_mat)
post_kappa <- colMeans(kappa_mat)
names(post_kappa) <- beta_cols_only
beta_ci <- apply(beta_mat, 2, quantile, c(0.025, 0.975))

term_labels <- c(
  "beta[1]" = "Seat comfort", "beta[2]" = "Dep/Arr time convenient", "beta[3]" = "Food and drink",
  "beta[4]" = "Gate location", "beta[5]" = "Inflight wifi service", "beta[6]" = "Inflight entertainment",
  "beta[7]" = "Online support", "beta[8]" = "Ease of Online booking", "beta[9]" = "On-board service",
  "beta[10]" = "Leg room service", "beta[11]" = "Baggage handling", "beta[12]" = "Checkin service",
  "beta[13]" = "Cleanliness", "beta[14]" = "Online boarding", "beta[15]" = "Age", "beta[16]" = "Flight Distance",
  "beta[17]" = "Class: Eco Plus", "beta[18]" = "Class: Business", "beta[19]" = "log(1+Departure Delay)",
  "beta[20]" = "Personal Travel", "beta[21]" = "disloyal Customer", "beta_arr" = "Mid-flight delay"
)

kappa_tbl <- data.frame(
  term = term_labels[beta_cols_only],
  kappa = round(post_kappa[beta_cols_only], 3),
  beta_mean = round(post_beta[beta_cols_only], 3),
  beta_lo = round(beta_ci[1, beta_cols_only], 3),
  beta_hi = round(beta_ci[2, beta_cols_only], 3)
) %>% arrange(kappa)

knitr::kable(kappa_tbl, row.names = FALSE,
             caption = "kappa_j near 0 = signal (unshrunk); near 1 = shrunk towards noise")
```

```{r kappa-plot, fig.width=8, fig.height=6}
# plot shrinkage coefficients k

kappa_df <- as.data.frame(kappa_mat)
colnames(kappa_df) <- term_labels[beta_cols_only]
kappa_long <- pivot_longer(kappa_df, cols = everything(), names_to = "Predictor", values_to = "Kappa")
kappa_long$Predictor <- factor(kappa_long$Predictor,
                               levels = kappa_tbl$term[order(-kappa_tbl$kappa)])

ggplot(kappa_long, aes(x = Predictor, y = Kappa)) +
  geom_violin(fill = "steelblue", alpha = 0.7, scale = "width") +
  geom_hline(yintercept = 0.5, color = "red", linetype = "dashed") +
  coord_flip() +
  labs(title = "Horseshoe shrinkage weights",
       subtitle = expression("near 0 (right) = signal preserved; near 1 (left) = shrunk towards zero"),
       y = expression(kappa[j]), x = NULL) +
  theme_minimal(base_size = 10)
```

```{r beta-plot, fig.width=7, fig.height=6}
kappa_tbl %>%
  mutate(term = factor(term, levels = term[order(kappa)])) %>%
  ggplot(aes(x = beta_mean, y = term, color = kappa < 0.5)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = beta_lo, xmax = beta_hi), height = 0.2) +
  geom_point(size = 2.2) +
  scale_color_manual(values = c("TRUE" = "#185FA5", "FALSE" = "grey60"), guide = "none") +
  labs(title = "Posterior coefficients under the horseshoe prior",
       subtitle = "Blue = kappa < 0.5 (signal); grey = kappa >= 0.5 (shrunk)",
       x = expression(beta[j]), y = NULL) +
  theme_minimal(base_size = 10)
```

```{r}
# store horseshoe MCMC for later analyses

beta0_summary <- c(mean = mean(sm[, "beta0"]),
                   quantile(sm[, "beta0"], c(0.025, 0.975)))
tau_summary   <- c(mean = mean(sm[, "tau"]),
                   quantile(sm[, "tau"],   c(0.025, 0.975)))
saveRDS(list(
  post_kappa  = post_kappa, 
  post_beta   = post_beta, 
  beta_ci     = beta_ci,
  beta0       = beta0_summary,
  tau         = tau_summary,
  term_labels = term_labels, 
  beta_cols   = beta_cols_only,
  lambda_cols = lambda_cols_only
  ), "../data/03_horseshoe_summary.rds")
```

## 3. Conlusions

**The "booking/connectivity" cluster visibly competes.** `Ease of Online booking` ($\kappa=$ `r kappa_tbl$kappa[kappa_tbl$term=="Ease of Online booking"]`) and `Online support` ($\kappa=$ `r kappa_tbl$kappa[kappa_tbl$term=="Online support"]`) are preserved noticeably more than the two other members of the same correlated cluster from Section 00, `Online boarding` ($\kappa=$ `r kappa_tbl$kappa[kappa_tbl$term=="Online boarding"]`) and `Inflight wifi service` ($\kappa=$ `r kappa_tbl$kappa[kappa_tbl$term=="Inflight wifi service"]`, the single most shrunk item in the whole model). This is exactly the mechanism the horseshoe was chosen for: four mutually correlated items compete, and the prior lets two of them carry the signal while treating the other two as largely redundant, instead of us picking one by hand.

**`Class: Business` shrinks once service quality is in the model.** In Section 02, Business class alone carried an odds ratio of about 4.1. Here, with all 14 service ratings included, its coefficient drops to `r kappa_tbl$beta_mean[kappa_tbl$term=="Class: Business"]` ($\kappa=$ `r kappa_tbl$kappa[kappa_tbl$term=="Class: Business"]`, credible interval crossing zero). The natural reading is mediation, not disappearance: Business passengers are more satisfied largely *because* they experience better seat comfort, entertainment, and service -- `Class` was partly standing in for those ratings in Section 02 simply because they weren't in the model yet.

**`Personal Travel` gets *stronger*, not weaker.** Section 02 found this covariate's effect nearly explained away by `Class`. Here, once the service ratings are controlling for the *quality* of the trip experience, `Personal Travel` re-emerges as one of the strongest and most confidently negative effects in the model ($\kappa=$ `r kappa_tbl$kappa[kappa_tbl$term=="Personal Travel"]`, credible interval `r kappa_tbl$beta_lo[kappa_tbl$term=="Personal Travel"]` to `r kappa_tbl$beta_hi[kappa_tbl$term=="Personal Travel"]`, entirely below zero). Section 02's finding and this one are not contradictory: they show that the earlier attenuation was specifically about `Class` soaking up part of the travel-type signal, not about travel type being unimportant.

A smaller, more cautionary finding: `Departure/Arrival time convenient` has a *negative* posterior mean (`r kappa_tbl$beta_mean[kappa_tbl$term=="Dep/Arr time convenient"]`, credible interval excluding zero) -- higher ratings associated with *lower* satisfaction. Taken at face value this is counter-intuitive. Recall from Section 00 that this item has one of the largest shares of structural zeros and sits in its own correlated cluster with `Seat comfort`, `Food and drink` and `Gate location`; we flagged, but did not yet resolve, the possibility that its zeros encode "not applicable" rather than genuine dissatisfaction. This coefficient is a candidate symptom of that ambiguity rather than a reliable substantive finding, and we revisit it directly in Section 05's structural-zero robustness check.

`disloyal Customer` remains, by a wide margin, the single strongest and most confidently preserved effect in the entire model ($\kappa=$ `r kappa_tbl$kappa[kappa_tbl$term=="disloyal Customer"]`) -- consistent across Sections 01, 02 and 03 regardless of how much else is in the model.

## 4. Results storage

```{r}
loglik_mat <- sm[, grepl("^log_lik", colnames(sm))]
loo_final <- loo::loo(loglik_mat)
waic_final <- loo::waic(loglik_mat)

beta_draws <- sm[, c("beta0", beta_cols_only)]

service_means <- colMeans(train[, service_items])
service_sds   <- apply(train[, service_items], 2, sd)

standardization <- list(
  service_means = service_means, service_sds = service_sds,
  Age_mean = mean(train$age), Age_sd = sd(train$age),
  FlightDist_mean = mean(train$flight_distance), FlightDist_sd = sd(train$flight_distance),
  logDepDelay_mean = mean(train$log_dep_delay), logDepDelay_sd = sd(train$log_dep_delay),
  logmid_mean = mean(train$log_mid_delay), logmid_sd = sd(train$log_mid_delay)
)

saveRDS(list(
  loo = loo_final, 
  waic = waic_final, 
  kappa_tbl = kappa_tbl,
  beta_draws = beta_draws, 
  standardization = standardization,
  service_items = service_items), "../data/03_horseshoe_final.rds")
```
