---
title: "04 - Prior Sensitivity and Structural Robustness"
subtitle: "Airline Customer Satisfaction -- Bayesian Learning & Monte Carlo Simulations, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
# loo_compare() is used by the model-comparison chunk below
library(loo)

rm(list = ls())
```

This notebook stress-tests the conclusions of Section 02 (vague-prior
logit) and Section 03 (horseshoe) against the choices we made along the
way. We ask three questions: does the vague model's posterior depend on
the prior *variance* we picked? Does the horseshoe's signal/noise
verdict depend on the *global scale* of its half-Cauchy prior? And does
the counter-intuitive `Departure/Arrival time convenient` coefficient
from Section 03 survive dropping the ambiguous structural-zero rows?

All the following checks require heacy JAGS fits, so we decided to split
those in separate script files (`04a_prior_sweep.R`,
`04b_horseshoe_prior_sweep.R`, `04c_structural_zero_refit.R`) to avoid excessive
long cells in this notebook. Scripts are meant to be executed ocne and
then re-use their stored output.

```{r load-data}
train <- read.csv("../data/airline_train_n1200.csv", stringsAsFactors = FALSE)
train$y <- as.integer(train$satisfaction == "satisfied")
std <- function(x) as.numeric(scale(x))
```

# 1. Logit model prior sensitivity

We refit Section 02's model (vague priors, dummy `Class` encoding) four
times, varying only the prior variance on every $\beta_j$:
$\sigma^2_\beta \in \{1, 10, 100, 10000\}$. The fits are produced by
`04a_prior_sweep.R`.

```{r vague-term-labels}
# Coefficient order returned by the sweep: beta[1..8], beta0, beta_arr.
term_labels_vague <- c("beta0" = "Intercept", "beta[1]" = "Age", "beta[2]" = "Seat comfort",
                       "beta[3]" = "Flight Distance", "beta[4]" = "Class: Eco Plus",
                       "beta[5]" = "Class: Business", "beta[6]" = "log(1+Departure Delay)",
                       "beta[7]" = "Personal Travel", "beta[8]" = "disloyal Customer",
                       "beta_arr" = "Mid-flight delay (signed-log)")
```

```{r prior-sweep-load}
sweep_results <- readRDS("../data/04a_prior_sweep.rds")
```

```{r prior-sweep-plot, fig.width=8, fig.height=6}
sweep_df <- bind_rows(lapply(names(sweep_results), function(v) {
  df <- sweep_results[[v]]
  df$term <- rownames(df)
  df$variance <- v
  df
}))
sweep_df$term_label <- term_labels_vague[sweep_df$term]
sweep_df$variance <- factor(sweep_df$variance,
                            levels = c("1", "10", "100", "10000"))

ggplot(sweep_df, aes(x = mean, y = term_label, color = variance)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.2,
                 position = position_dodge(width = 0.6)) +
  geom_point(position = position_dodge(width = 0.6), size = 2) +
  labs(title = "Posterior coefficients across four prior variances",
       subtitle = expression(sigma[beta]^2 %in% "{1, 10, 100, 10000}"),
       x = expression(beta), y = NULL, color = expression(sigma[beta]^2)) +
  theme_minimal(base_size = 10)
```

```{r prior-sweep-range}
range_tbl <- sweep_df %>%
  group_by(term_label) %>%
  summarise(mean_range = max(mean) - min(mean), .groups = "drop") %>%
  arrange(desc(mean_range))
knitr::kable(range_tbl, digits = 4,
             caption = "Range of posterior means across the four prior variances")
```

The largest movement across four orders of magnitude of prior variance
is only `r sprintf("%.3f", max(range_tbl$mean_range))` on the log-odds
scale, for a coefficient whose posterior mean is itself on the order of
1-2. At $n=$ `r nrow(train)`. So th likelihood dominates the tested priors.

# Horseshoe prior sensitivity: does the global scale matter?

The horseshoe's shrinkage behaviour is governed by the scale $s$ of the 
half-Cauchy prior put on the **global** scale $\tau$: 
$$\tau \sim \mathcal{C}^+(0, s)$$
Small $s$ pulls every coefficient hard towards zero (more aggressive sparsity),
large $s$ relaxes the global shrinkage. 
This is the parameter whose choice could plausibly change *which* covariates the 
model keeps, so it is the one worth testing.

`04b_horseshoe_prior_sweep.R` refits the exact Section 03 model three times,
varying only this global scale, $s \in \{0.1, 1, 10\}$ ($s = 1$ is the
Section 03 default), while the local scales $\lambda_j$ keep the
standard $\mathcal{C}^+(0,1)$. For each fit it stores the posterior mean
and $95\%$ credible interval of every coefficient, and the per-predictor
shrinkage weight $\kappa_j = 1/(1+\lambda_j^2)$ (near $0$ = signal, near
$1$ = shrunk, exactly as in Section 03).

```{r hs-sweep-load}
hs_sweep <- readRDS("../data/04b_horseshoe_sweep.rds")
hs_scales <- names(hs_sweep)

hs_df <- bind_rows(lapply(hs_scales, function(s) {
  d <- hs_sweep[[s]]
  d$scale <- s
  d
}))
hs_df$scale <- factor(hs_df$scale, levels = hs_scales)
```

## Coefficients across global scales

```{r hs-sweep-beta-plot, fig.width=9, fig.height=8}
ggplot(hs_df, aes(x = beta_mean, y = reorder(term, beta_mean), color = scale)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = beta_lo, xmax = beta_hi), height = 0.25,
                 position = position_dodge(width = 0.6)) +
  geom_point(position = position_dodge(width = 0.6), size = 1.8) +
  labs(title = "Horseshoe posterior coefficients across three global scales",
       subtitle = expression(tau %~% "C+(0, s), s in {0.1, 1, 10}"),
       x = expression(beta[j]), y = NULL, color = "global scale s") +
  theme_minimal(base_size = 10)
```

```{r hs-sweep-beta-range}
hs_range <- hs_df %>%
  group_by(term) %>%
  summarise(beta_range = max(beta_mean) - min(beta_mean), .groups = "drop") %>%
  arrange(desc(beta_range))
knitr::kable(head(hs_range, 8), digits = 3,
             caption = "Largest movement of the posterior mean across the three global scales")
```

## Shrinkage weights and signal/noise stability

```{r hs-sweep-kappa-wide}
kappa_wide <- hs_df %>%
  select(term, scale, kappa) %>%
  pivot_wider(names_from = scale, values_from = kappa)
knitr::kable(kappa_wide, digits = 3,
             caption = "Posterior shrinkage weight kappa_j at each global scale")
```

```{r hs-sweep-signal-stability}
signal_wide <- hs_df %>%
  mutate(signal = kappa < 0.5) %>%
  select(term, scale, signal) %>%
  pivot_wider(names_from = scale, values_from = signal)

signal_cols <- setdiff(names(signal_wide), "term")
signal_wide$stable <- apply(signal_wide[, signal_cols], 1,
                            function(r) length(unique(r)) == 1)
n_flips <- sum(!signal_wide$stable)

# predictors classified as signal (kappa < 0.5) under the default scale s = 1
default_signal <- signal_wide$term[signal_wide[["1"]]]

knitr::kable(signal_wide, caption = "Signal (kappa < 0.5) at each scale; `stable` = same verdict across all three")
```

Across a hundred-fold range of the global scale, `r n_flips` of the 22
predictors change their signal/noise classification. The covariates the
model calls signal at the default scale ($s = 1$) --
`r paste(default_signal, collapse = ", ")` -- are
`r if (n_flips == 0) "exactly the set it keeps at every scale" else "largely preserved across scales"`,
and the largest shift in any posterior mean is
`r sprintf("%.3f", max(hs_range$beta_range))` on the log-odds scale. The
strong effects identified in Section 03 (`disloyal Customer`,
`Personal Travel`, `Seat comfort`) stay strong and confidently non-zero
regardless of $s$; only the coefficients already sitting near the
$\kappa = 0.5$ boundary drift, which is exactly where sensitivity to the
global scale is expected and where we drew no firm conclusions in the
first place. The headline reading of Section 03 is therefore not an
artefact of the particular half-Cauchy scale we chose.



# Structural-zero robustness check

Section 03 speculated that `Departure/Arrival time convenient`'s
counter-intuitive *negative* coefficient might be an artefact of the
structural zeros flagged in Section 00 (zeros that plausibly mean "not
applicable" rather than genuine dissatisfaction). We test that directly
by refitting the full horseshoe model with those rows excluded
(`04c_structural_zero_refit.R`).

```{r structural-zero-load}
# Section 03 horseshoe summary (full data). The saved object holds the pieces,
# not an assembled table, so rebuild one with the same columns as the refit's
# result_tbl, otherwise the two cannot be bound together below.
s3   <- readRDS("../data/03_horseshoe_summary.rds")
s3$kappa_tbl <- data.frame(
  term      = s3$term_labels[s3$beta_cols],
  kappa     = as.numeric(s3$post_kappa[s3$beta_cols]),
  beta_mean = as.numeric(s3$post_beta[s3$beta_cols]),
  beta_lo   = as.numeric(s3$beta_ci[1, s3$beta_cols]),
  beta_hi   = as.numeric(s3$beta_ci[2, s3$beta_cols]),
  row.names = NULL)
zref <- readRDS("../data/04c_structural_zero_refit.rds")
cat("Rows excluded:", zref$n_excluded, "of", zref$n_excluded + zref$n_remaining, "\n")
cat("Rows remaining:", zref$n_remaining, "\n")
```

```{r structural-zero-table}
knitr::kable(zref$result_tbl, row.names = FALSE, digits = 3,
             caption = "Horseshoe refit excluding structural-zero rows")
```

```{r structural-zero-compare, fig.width=7, fig.height=5}
full_tbl <- s3$kappa_tbl %>% mutate(fit = "Full data")
zero_tbl <- zref$result_tbl %>% mutate(fit = "Excluding structural zeros")

compare_df <- bind_rows(full_tbl, zero_tbl) %>%
  filter(term %in% c("Dep/Arr time convenient", "Seat comfort", "Food and drink", "Gate location"))

ggplot(compare_df, aes(x = beta_mean, y = term, color = fit)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = beta_lo, xmax = beta_hi), height = 0.15,
                 position = position_dodge(width = 0.4)) +
  geom_point(position = position_dodge(width = 0.4), size = 2.2) +
  labs(title = "The 'seat/catering/ground' cluster: full data vs. excluding structural zeros",
       x = expression(beta), y = NULL, color = NULL) +
  theme_minimal(base_size = 10)
```

**The hypothesis does not survive the check.**
`Departure/Arrival time convenient`'s negative coefficient does not
shrink towards zero when the ambiguous rows are removed -- it gets
*more* negative and more confident (from
`r sprintf("%.2f", full_tbl$beta_mean[full_tbl$term=="Dep/Arr time convenient"])`
to
`r sprintf("%.2f", zero_tbl$beta_mean[zero_tbl$term=="Dep/Arr time convenient"])`).
Whatever drives this effect, it is not simply the structural-zero
ambiguity we suspected in Section 03; we report this as an open,
unresolved finding rather than force a tidier story onto it. On the
other hand, `Seat comfort` and `Food and drink` -- the other two members
of that same correlated cluster -- both strengthen substantially once
the ambiguous rows are excluded (`Seat comfort` $\kappa$ drops from
`r full_tbl$kappa[full_tbl$term=="Seat comfort"]` to
`r zero_tbl$kappa[zero_tbl$term=="Seat comfort"]`), consistent with
those zeros adding genuine noise to *those* two items specifically.

```{r}
## Model comparison: WAIC and PSIS-LOO

s2 <- readRDS("../data/02_logit_mcmc.rds")
s3 <- readRDS("../data/03_horseshoe_final.rds")

print(s2$loo)
print(s3$loo)

loo_cmp <- loo_compare(list(logit = s2$loo, horseshoe = s3$loo))
print(loo_cmp)
```

```{r}
loo_tbl <- data.frame(
  Model = c("Vague-prior logit", "Horseshoe"),
  elpd_loo = c(s2$loo$estimates["elpd_loo","Estimate"], s3$loo$estimates["elpd_loo","Estimate"]),
  SE       = c(s2$loo$estimates["elpd_loo","SE"],       s3$loo$estimates["elpd_loo","SE"]),
  p_loo    = c(s2$loo$estimates["p_loo","Estimate"],    s3$loo$estimates["p_loo","Estimate"]),
  looic    = c(s2$loo$estimates["looic","Estimate"],    s3$loo$estimates["looic","Estimate"])
)
knitr::kable(round(loo_tbl[,-1], 1), row.names = FALSE)
```

