---
title: "06 - Prediction and Validation"
subtitle: "Airline Customer Satisfaction -- Bayesian Learning & Monte Carlo Simulations, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                       fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
```

Prediction here is pure matrix algebra on already-fitted posterior draws
-- no new MCMC sampling -- so it stays cheap even against the full
held-out test set. We evaluate the two JAGS models from Sections 02 and
03 against the untouched test set, which neither model has seen in any
form since the Section 00 split.

```{r load}
test <- read.csv("../data/airline_test_holdout.csv", stringsAsFactors = FALSE)
test$y <- as.integer(test$satisfaction == "satisfied")
n_test <- nrow(test)

s2 <- readRDS("../data/02_logit_mcmc.rds")
s3 <- readRDS("../data/03_horseshoe_final.rds")

invlogit <- function(x) 1 / (1 + exp(-x))
```

Test set size: `r format(n_test, big.mark = ",")` rows.

# Preparing the test-set covariates

Every continuous covariate is standardized using the **training-set**
mean and SD saved by Sections 02 and 03 -- never statistics recomputed
on the test set, which would leak test information into preprocessing.

```{r test-covariates}
std_with <- function(x, mean, sd) (x - mean) / sd

logDepDelay_test  <- log1p(test$dep_delay)
signedLogArr_test <- test$log_mid_delay

Class_EcoPlus_test  <- as.integer(test$class == "Eco Plus")
Class_Business_test <- as.integer(test$class == "Business")
TypeTravel_Personal_test <- as.integer(test$travel_type == "Personal Travel")
CustType_disloyal_test   <- as.integer(test$customer_type == "disloyal Customer")

n_missing_test <- sum(is.na(signedLogArr_test))
```

`r n_missing_test` test rows (`r sprintf("%.2f%%", 100*n_missing_test/n_test)`)
are missing the arrival-delay residual. JAGS's stochastic-imputation
trick only applies during fitting; for out-of-sample prediction on new
customers, we set the standardized residual to its assumed prior mean of
0 for these rows -- the posterior-predictive-consistent choice, since
both models' own `arr_delay[i] ~ dnorm(0,1)` prior says exactly that this
is the expected value in the absence of other information.

```{r test-covariates-2}
# --- Section 02 design (dummy Class, 8 covariates) ---
X_test_s2 <- cbind(
  Age_std        = std_with(test$age, s2$standardization$Age_mean, s2$standardization$Age_sd),
  Seat_std       = std_with(test$seat_comfort, s2$standardization$Seat_mean, s2$standardization$Seat_sd),
  FlightDist_std = std_with(test$flight_distance, s2$standardization$FlightDist_mean, s2$standardization$FlightDist_sd),
  Class_EcoPlus  = Class_EcoPlus_test,
  Class_Business = Class_Business_test,
  logDepDelay_std = std_with(logDepDelay_test, s2$standardization$logdep_mean, s2$standardization$logDepDelay_sd),
  TypeTravel_Personal = TypeTravel_Personal_test,
  CustType_disloyal   = CustType_disloyal_test
)
arr_delay_test_s2 <- std_with(signedLogArr_test, s2$standardization$logmid_mean, s2$standardization$logmid_sd)
arr_delay_test_s2[is.na(arr_delay_test_s2)] <- 0

# --- Section 03 design (14 service items + covariates) ---
service_items <- s3$service_items
X_service_test <- sapply(service_items, function(nm) {
  std_with(test[[nm]], s3$standardization$service_means[nm], s3$standardization$service_sds[nm])
})
X_extra_test <- cbind(
  Age_std        = std_with(test$age, s3$standardization$Age_mean, s3$standardization$Age_sd),
  FlightDist_std = std_with(test$flight_distance, s3$standardization$FlightDist_mean, s3$standardization$FlightDist_sd),
  Class_EcoPlus  = Class_EcoPlus_test,
  Class_Business = Class_Business_test,
  logDepDelay_std = std_with(logDepDelay_test, s3$standardization$logDepDelay_mean, s3$standardization$logDepDelay_sd),
  TypeTravel_Personal = TypeTravel_Personal_test,
  CustType_disloyal   = CustType_disloyal_test
)
X_test_s3 <- cbind(X_service_test, X_extra_test)
arr_delay_test_s3 <- std_with(signedLogArr_test, s3$standardization$logmid_mean, s3$standardization$logmid_sd)
arr_delay_test_s3[is.na(arr_delay_test_s3)] <- 0
```

# Posterior-mean prediction

For every posterior draw $s$ and test customer $i$:
$p_i^{(s)} = \text{logit}^{-1}\!\left(\beta_0^{(s)} + x_i\beta^{(s)} + \beta_{\text{arr}}^{(s)}\,\text{arr\_delay}_i\right)$,
averaged over draws for the final prediction $\hat p_i$. We loop over
the (thinned, 1000-draw) posterior rather than materialising a
$`r format(n_test,big.mark=",")` \times 1000$ matrix, to keep memory
bounded -- and, in the same pass, draw a genuine posterior-predictive
replicate $y_i^{(s)} \sim \text{Bernoulli}(p_i^{(s)})$ per customer for
the subgroup check in a later section.

```{r predict-function}
predict_model <- function(beta_draws, X, arr_delay, group_vars = NULL) {
  S <- nrow(beta_draws)
  N <- nrow(X)
  beta_cols <- grep("^beta\\[", colnames(beta_draws), value = TRUE)
  beta_cols <- beta_cols[order(as.numeric(gsub("beta\\[|\\]", "", beta_cols)))]
  Bmat  <- beta_draws[, beta_cols, drop = FALSE]
  beta0 <- beta_draws[, "beta0"]
  barr  <- beta_draws[, "beta_arr"]

  p_sum <- numeric(N)
  group_rep <- if (!is.null(group_vars)) {
    lapply(group_vars, function(g) matrix(NA_real_, nrow = S, ncol = length(unique(g)),
                                            dimnames = list(NULL, sort(unique(g)))))
  } else NULL
  names(group_rep) <- names(group_vars)

  for (s in seq_len(S)) {
    lin <- beta0[s] + as.vector(X %*% Bmat[s, ]) + barr[s] * arr_delay
    p <- invlogit(lin)
    p_sum <- p_sum + p
    if (!is.null(group_vars)) {
      y_rep <- rbinom(N, 1, p)
      for (gname in names(group_vars)) {
        g <- group_vars[[gname]]
        rates <- tapply(y_rep, g, mean)
        group_rep[[gname]][s, names(rates)] <- rates
      }
    }
  }
  list(p_mean = p_sum / S, group_rep = group_rep)
}

set.seed(123)
pred_s2 <- predict_model(s2$beta_draws, X_test_s2, arr_delay_test_s2)
pred_s3 <- predict_model(s3$beta_draws, X_test_s3, arr_delay_test_s3,
                          group_vars = list(Class = test$class, CustomerType = test$customer_type,
                                             TypeOfTravel = test$travel_type))

p_s2 <- pred_s2$p_mean
p_s3 <- pred_s3$p_mean
```

# Did the horseshoe's complexity earn its keep out-of-sample?

Section 05 showed the horseshoe fits much better in-sample and by
PSIS-LOO. Here we check the more skeptical question directly: does that
translate into better held-out predictions on customers neither model
has ever seen? We also bring in Section 04's BAS model here: BAS's
`predict()` method supports genuine BMA prediction on new data (no
separate coefficient-draw export needed, unlike JAGS), so the same
question can be asked of all three model families side by side.

```{r bas-predict}
suppressMessages(library(BAS))
fit_bas <- readRDS("../data/A01_BAS_model.rds")

df_test_bas <- data.frame(
  X_service_test,
  Age_std = X_extra_test[, "Age_std"], FlightDist_std = X_extra_test[, "FlightDist_std"],
  Class_EcoPlus = X_extra_test[, "Class_EcoPlus"], Class_Business = X_extra_test[, "Class_Business"],
  # Column names must match fit_bas$namesx exactly: BAS rebuilds the model
  # frame from the formula, so a renamed column is simply a missing variable.
  logdep_std = X_extra_test[, "logDepDelay_std"],
  TypeTravel_Personal = X_extra_test[, "TypeTravel_Personal"],
  CustType_disloyal = X_extra_test[, "CustType_disloyal"],
  logmid_delay = arr_delay_test_s3
)

# BAS's predict(..., estimator="BMA") builds an internal n_rows x n_models
# matrix (here ~1,800 visited models); at the full ~129k-row test set that
# is >1.5GB, which was enough to get the R process OOM-killed.
# Predicting in batches keeps peak memory bounded to a few hundred MB.
predict_bas_batched <- function(fit, newdata, batch_size = 10000) {
  n <- nrow(newdata)
  out <- numeric(n)
  for (st in seq(1, n, by = batch_size)) {
    en <- min(st + batch_size - 1, n)
    out[st:en] <- predict(fit, newdata = newdata[st:en, ], type = "response", estimator = "BMA")$fit
    invisible(gc(full = TRUE))
  }
  out
}

p_bas <- predict_bas_batched(fit_bas, df_test_bas)
```

```{r auc-function}
auc_mw <- function(probs, labels) {
  r <- rank(probs)
  n1 <- as.numeric(sum(labels == 1)); n0 <- as.numeric(sum(labels == 0))
  (sum(r[labels == 1]) - n1 * (n1 + 1) / 2) / (n1 * n0)
}
brier <- function(probs, labels) mean((probs - labels)^2)

model_cmp <- data.frame(
  model = c("Section 02 (vague, 9 predictors)", "Section 03 (horseshoe, 22 predictors)",
            "Section 04 (BAS, BMA over model space)"),
  AUC   = c(auc_mw(p_s2, test$y), auc_mw(p_s3, test$y), auc_mw(p_bas, test$y)),
  Brier = c(brier(p_s2, test$y), brier(p_s3, test$y), brier(p_bas, test$y))
)
knitr::kable(model_cmp, digits = 4)
```

```{r roc-3way, fig.width=6.5, fig.height=5.5}
roc_s2  <- compute_roc_curve <- function(probs, labels, n_thresh = 100) {
  thresholds <- seq(0, 1, length.out = n_thresh)
  sapply(thresholds, function(t) {
    pred <- as.integer(probs >= t)
    tp <- sum(pred == 1 & labels == 1); fn <- sum(pred == 0 & labels == 1)
    fp <- sum(pred == 1 & labels == 0); tn <- sum(pred == 0 & labels == 0)
    c(tpr = tp / (tp + fn), fpr = fp / (fp + tn))
  }) %>% t() %>% as.data.frame()
}

roc_all <- bind_rows(
  compute_roc_curve(p_s2, test$y) %>% mutate(model = sprintf("Section 02 (AUC=%.3f)", model_cmp$AUC[1])),
  compute_roc_curve(p_s3, test$y) %>% mutate(model = sprintf("Section 03 (AUC=%.3f)", model_cmp$AUC[2])),
  compute_roc_curve(p_bas, test$y) %>% mutate(model = sprintf("Section 04 BAS (AUC=%.3f)", model_cmp$AUC[3]))
)

ggplot(roc_all, aes(fpr, tpr, color = model)) +
  geom_line(linewidth = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  labs(title = "Test-set ROC: all three model families", x = "False positive rate", y = "True positive rate",
       color = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom")
```

The horseshoe model wins on both metrics out-of-sample, not just on
PSIS-LOO -- the extra 13 service-rating predictors and the shrinkage
machinery are not overfitting to the training subsample; they carry
real, generalisable signal. **BAS lands close behind the horseshoe** on
both AUC and Brier, well ahead of the simpler vague-prior model and
clearly not overfitting either, despite Section 04's very different
estimation strategy (discrete model averaging over a stochastically
searched model space, rather than a single model with continuous
shrinkage). This is the strongest form of the cross-check Section 04 set
out to do: two structurally different Bayesian approaches not only
agree on *which* variables matter (95% agreement, Section 04) but
deliver near-identical *predictive* performance on data neither has
seen. We use the horseshoe (`p_s3`) for the model-specific deep dives
below, since it is the single best performer, but the BAS numbers here
are the real payoff of Section 04's cross-check.

# ROC curve and confusion matrices

```{r roc-curve, fig.width=6, fig.height=5}
compute_roc <- function(probs, labels, n_thresh = 200) {
  thresholds <- seq(0, 1, length.out = n_thresh)
  sapply(thresholds, function(t) {
    pred <- as.integer(probs >= t)
    tp <- sum(pred == 1 & labels == 1); fn <- sum(pred == 0 & labels == 1)
    fp <- sum(pred == 1 & labels == 0); tn <- sum(pred == 0 & labels == 0)
    c(threshold = t, tpr = tp / (tp + fn), fpr = fp / (fp + tn))
  }) %>% t() %>% as.data.frame()
}

roc_df <- compute_roc(p_s3, test$y)
auc_val <- auc_mw(p_s3, test$y)

ggplot(roc_df, aes(fpr, tpr)) +
  geom_line(color = "#185FA5", linewidth = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  labs(title = sprintf("Test-set ROC (AUC = %.3f)", auc_val),
       x = "False positive rate", y = "True positive rate") +
  theme_minimal(base_size = 11)
```

```{r confusion-matrices}
make_confusion <- function(probs, labels, threshold) {
  pred <- as.integer(probs >= threshold)
  tab <- table(Predicted = pred, Actual = labels)
  sens <- tab["1","1"] / sum(tab[,"1"])
  spec <- tab["0","0"] / sum(tab[,"0"])
  acc  <- sum(diag(tab)) / sum(tab)
  list(table = tab, sensitivity = sens, specificity = spec, accuracy = acc)
}

base_rate <- mean(read.csv("../data/airline_train_n1200.csv", stringsAsFactors = FALSE)$satisfaction == "satisfied")

cm_05   <- make_confusion(p_s3, test$y, 0.5)
cm_base <- make_confusion(p_s3, test$y, base_rate)

cat("--- Threshold = 0.5 ---\n"); print(cm_05$table)
cat(sprintf("Sensitivity: %.3f  Specificity: %.3f  Accuracy: %.3f\n\n",
            cm_05$sensitivity, cm_05$specificity, cm_05$accuracy))

cat(sprintf("--- Threshold = base rate (%.3f) ---\n", base_rate)); print(cm_base$table)
cat(sprintf("Sensitivity: %.3f  Specificity: %.3f  Accuracy: %.3f\n",
            cm_base$sensitivity, cm_base$specificity, cm_base$accuracy))
```

As usual with a threshold near 0.5 on a moderately balanced outcome,
sensitivity and specificity trade off; shifting to the empirical base
rate re-balances them without needing a full loss-function analysis --
which we do properly in the last section below.

# Calibration and the Brier score

```{r calibration, fig.width=6, fig.height=5}
calib_df <- data.frame(p = p_s3, y = test$y) %>%
  mutate(decile = ntile(p, 10)) %>%
  group_by(decile) %>%
  summarise(mean_predicted = mean(p), observed_rate = mean(y), n = n(), .groups = "drop")

knitr::kable(calib_df, digits = 3, caption = "Calibration by decile of predicted probability")

ggplot(calib_df, aes(mean_predicted, observed_rate)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_point(size = 2.5, color = "#185FA5") +
  geom_line(color = "#185FA5") +
  coord_equal(xlim = c(0,1), ylim = c(0,1)) +
  labs(title = "Calibration plot: mean predicted vs. observed satisfaction rate",
       x = "Mean predicted probability (decile)", y = "Observed satisfaction rate") +
  theme_minimal(base_size = 11)

cat(sprintf("Brier score (horseshoe): %.4f\n", brier(p_s3, test$y)))
```

Points sit close to the diagonal across all ten deciles: the model's
stated confidence matches reality reasonably well, not just its
discrimination (AUC). A Brier score of
`r sprintf("%.4f", brier(p_s3, test$y))` is a genuine improvement over
the naive always-predict-the-base-rate Brier score of
`r sprintf("%.4f", brier(base_rate, test$y))`.

# Posterior predictive check by subgroup

Rather than stopping at point predictions, we simulate replicated
satisfaction outcomes from the posterior draws and compare the
*simulated* subgroup satisfaction rate to what we actually observe in
the test set -- the same three subgroups Section 01 analysed, now
checked against the model's out-of-sample predictions on
`r format(n_test, big.mark=",")` customers it has never seen.

```{r ppc-plot, fig.width=9, fig.height=3.5}
observed_rates <- list(
  Class = tapply(test$y, test$class, mean),
  CustomerType = tapply(test$y, test$customer_type, mean),
  TypeOfTravel = tapply(test$y, test$travel_type, mean)
)

ppc_df <- bind_rows(lapply(names(pred_s3$group_rep), function(gname) {
  mat <- pred_s3$group_rep[[gname]]
  as.data.frame(mat) %>%
    pivot_longer(everything(), names_to = "level", values_to = "simulated_rate") %>%
    mutate(facet = gname)
}))

obs_df <- bind_rows(lapply(names(observed_rates), function(gname) {
  data.frame(level = names(observed_rates[[gname]]), observed = as.numeric(observed_rates[[gname]]),
             facet = gname)
}))

p_ppc <- ggplot(ppc_df, aes(simulated_rate, fill = level)) +
  geom_density(alpha = 0.5) +
  geom_vline(data = obs_df, aes(xintercept = observed, color = level), linetype = "dashed", linewidth = 0.9) +
  facet_wrap(~ facet, scales = "free") +
  labs(title = "Posterior predictive check: simulated vs. observed subgroup satisfaction rate",
       subtitle = "Dashed lines = actually observed rate in the held-out test set",
       x = "Simulated satisfaction rate", y = "Density", fill = NULL, color = NULL) +
  theme_minimal(base_size = 9)
dir.create("figs", showWarnings = FALSE)
dir.create("../figs", showWarnings = FALSE)
ggsave("../figs/ppc_subgroups.png", p_ppc, width = 9, height = 3.5, dpi = 150)
p_ppc
```

The observed rates (dashed lines) fall inside the bulk of each
simulated posterior predictive distribution -- the model is not
systematically over- or under-predicting satisfaction for any Class,
Customer Type, or Type of Travel subgroup on genuinely unseen data. This
is the real posterior predictive check the roadmap called for, not just
a point-prediction accuracy number.

# Decision threshold via an explicit loss function

A plain 0.5 or ROC-optimal cutoff treats both error types as equally
costly. In practice, missing a recoverable dissatisfied customer (a lost
retention opportunity) is plausibly more expensive than a wasted
proactive-outreach contact to someone who was satisfied anyway. Flag a
customer as "at risk" when $\hat p_i < t$, and let:

- $C_{\text{miss}}$ = cost of failing to flag an actually-dissatisfied customer,
- $C_{\text{waste}}$ = cost of flagging an actually-satisfied customer.

Expected cost at threshold $t$: $C_{\text{miss}} \cdot P(y=1_{\text{sat}}, \hat p \ge t \mid y=0) + C_{\text{waste}} \cdot P(\hat p < t \mid y=1)$,
estimated empirically on the test set.

```{r loss-function, fig.width=7, fig.height=4.5}
expected_cost <- function(probs, labels, threshold, c_miss, c_waste) {
  flagged <- probs < threshold
  miss  <- sum(!flagged & labels == 0)   # dissatisfied, not flagged
  waste <- sum(flagged & labels == 1)    # satisfied, flagged anyway
  (c_miss * miss + c_waste * waste) / length(labels)
}

thresholds <- seq(0.05, 0.95, by = 0.01)
cost_ratios <- list("1:1 (equal cost)" = c(1, 1), "3:1" = c(3, 1), "5:1" = c(5, 1))

cost_df <- bind_rows(lapply(names(cost_ratios), function(nm) {
  cm <- cost_ratios[[nm]]
  data.frame(threshold = thresholds,
             cost = sapply(thresholds, expected_cost, probs = p_s3, labels = test$y,
                            c_miss = cm[1], c_waste = cm[2]),
             ratio = nm)
}))

optimal_tbl <- cost_df %>% group_by(ratio) %>% slice_min(cost, n = 1) %>%
  select(ratio, threshold, cost)

ggplot(cost_df, aes(threshold, cost, color = ratio)) +
  geom_line(linewidth = 1) +
  geom_point(data = optimal_tbl, size = 3) +
  labs(title = "Expected cost per customer vs. decision threshold",
       subtitle = "Points mark the cost-minimising threshold for each cost ratio",
       x = "Threshold t (flag as at-risk if predicted P(satisfied) < t)",
       y = "Expected cost per customer (arbitrary units)", color = "Cost ratio\n(miss:waste)") +
  theme_minimal(base_size = 10)

knitr::kable(optimal_tbl, digits = 3, caption = "Cost-minimising threshold by assumed cost ratio")
```

The optimal threshold moves from
`r sprintf("%.2f", optimal_tbl$threshold[optimal_tbl$ratio=="1:1 (equal cost)"])`
(equal costs, close to the base rate) up to
`r sprintf("%.2f", optimal_tbl$threshold[optimal_tbl$ratio=="5:1"])`
once missing a recoverable customer is assumed five times costlier than
a wasted contact -- exactly the direction intuition suggests, and a
genuinely different, business-relevant answer from either the 0.5 or
ROC-optimal defaults. We do not claim to know the airline's true cost
ratio; the point is that the *right* threshold is a business decision
informed by this curve, not a statistical default.
