---
title: "A01_BAS_comparison"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                       fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(BAS)

rm(list = ls())
```

After fitting our feature selection horseshoe prior-based model, we'd
like to compare it with BAS model in order to verify if the two methods
agree on which covariates are more useful in explaining the target
variable.

## 0. Dataset load

```{r load-data}
train <- read.csv("../data/airline_train_n1200.csv", stringsAsFactors = FALSE)
train$y <- as.integer(train$satisfaction == "satisfied")

std <- function(x) as.numeric(scale(x))

service_items <- c("seat_comfort", "time_convenient", "food_drink", "gate_location", "wifi", "entertainment", "online_support", "ease_booking", "onboard_service","legroom", "baggage", "checkin", "cleanliness", "online_boarding" )
X_service <- apply(train[, service_items], 2, std)

logdep_std      <- std(train$log_dep_delay)
Age_std         <- std(train$age)
FlightDist_std  <- std(train$flight_distance)
logmid_delay    <- std(train$log_mid_delay)

Class_EcoPlus  <- as.integer(train$cabin_class == "Eco Plus")
Class_Business <- as.integer(train$cabin_class == "Business")
TypeTravel_Personal <- as.integer(train$travel_type == "Personal Travel")
CustType_disloyal   <- as.integer(train$customer_type == "disloyal Customer")

X_extra <- cbind(Age_std, FlightDist_std, Class_EcoPlus, Class_Business,
                 logdep_std, TypeTravel_Personal, CustType_disloyal, logmid_delay)
X_full <- cbind(X_service, X_extra)

df_bas <- data.frame(y = train$y, X_full)
```

# 1. `bas.glm` fit with a binomial family

With `r ncol(X_full)+1` predictors, $2^{`r ncol(X_full)+1`} \approx$
`r format(round(2^(ncol(X_full)+1)/1e6,1), big.mark=",")` million
candidate models -- too many to enumerate exhaustively, so we use
`method = "MCMC"` in BAS to stochastically search the model space. Later
section will asses whether this search suffers from instability or not.

```{r fit-glm-g}
set.seed(123)
fit_glm_g <- bas.glm(y ~ ., data = df_bas, family = binomial(),
                      betaprior = g.prior(nrow(df_bas)), modelprior = uniform(),
                      method = "MCMC", MCMC.iterations = 100000)
```

```{r top-models}
knitr::kable(round(summary(fit_glm_g, n.models = 5), 3),
                          caption = "Top 5 models by posterior probability (g-prior)")
```

# 2. Parameter space convergence

Now we test the goodness of the MCMC search in the parameter space
executed by BAS.

### Convergence within one run.

BAS reports two independent estimates of each posterior inclusion
probability:

- `probne0.RB` : takes the set of distinct model visited containing the
  variable $j$, use their marginal likelihoods, renomalizes and sum the
  mass of those models

- `probne0.MCM`: it's just the raw fraction of iterations the chain
  spent in some model containing variable $j$ If the chain has
  converged, they should agree closely.

```{r diagnostics-plot, fig.width=5, fig.height=5}
diag_df <- data.frame(renormalized = fit_glm_g$probne0.RN,
                       mcmc = fit_glm_g$probne0.MCMC)
ggplot(diag_df, aes(renormalized, mcmc)) +
  geom_point(color = "#185FA5", alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  coord_equal(xlim = c(0,1), ylim = c(0,1)) +
  labs(title = "MCMC convergence check: PIP estimates should lie on the diagonal",
       x = "PIP (renormalized marginal likelihood)", y = "PIP (raw MCMC frequency)") +
  theme_minimal(base_size = 11)
```

Correlation between the two estimates is
`r sprintf("%.4f", cor(fit_glm_g$probne0.RN, fit_glm_g$probne0.MCMC))`,
maximum absolute difference
`r sprintf("%.3f", max(abs(fit_glm_g$probne0.RN - fit_glm_g$probne0.MCMC)))`
-- the points sit essentially on the diagonal, the sign this run has
converged.

### stability across seeds

We refit the identical model twice more with different random seeds and
compare posterior inclusion probabilities (PIP) to check whether they
agree or not.

```{r seed-stability, cache=TRUE}
run_seed <- function(seed) {
  set.seed(seed)
  fit <- bas.glm(y ~ ., data = df_bas, family = binomial(),
                 betaprior = g.prior(nrow(df_bas)), modelprior = uniform(),
                 method = "MCMC", MCMC.iterations = 100000)
  fit$probne0
}
pip_seed1 <- run_seed(1)
pip_seed2 <- run_seed(2)

pip_df <- data.frame(term = fit_glm_g$namesx,
                      seed123 = fit_glm_g$probne0,
                      seed1 = pip_seed1, seed2 = pip_seed2)
pip_df$range <- apply(pip_df[, 2:4], 1, function(x) max(x) - min(x))
```

```{r seed-stability-table}
knitr::kable(pip_df %>% arrange(desc(range)) %>% head(8), row.names = FALSE, digits = 3,
             caption = "Largest PIP swings across 3 seeds (of 22 predictors)")
cat(sprintf("Max PIP range across seeds: %.3f\nMean PIP range across seeds: %.4f\n",
            max(pip_df$range), mean(pip_df$range)))
```

### BIC prior as a robustness check

We try another prior on the beta coefficients since, if the search is
stable, the effect of the prior should vanish if the number of
iterations is high enough.

```{r bic-prior, cache=TRUE}
set.seed(123)
fit_glm_bic <- bas.glm(y ~ ., data = df_bas, family = binomial(),
                        betaprior = bic.prior(nrow(df_bas)), modelprior = uniform(),
                        method = "MCMC", MCMC.iterations = 100000)

cmp_prior <- data.frame(term = fit_glm_g$namesx, g_prior = round(fit_glm_g$probne0, 3),
                         BIC = round(fit_glm_bic$probne0, 3))
cmp_prior$abs_diff <- round(abs(cmp_prior$g_prior - cmp_prior$BIC), 3)
```

```{r bic-table}
knitr::kable(cmp_prior %>% arrange(desc(abs_diff)) %>% head(6), row.names = FALSE,
             caption = "Largest g-prior vs. BIC-prior PIP differences")
cat("Max |g-prior - BIC| difference:", max(cmp_prior$abs_diff), "\n")
```

With 100,000 MCMC iterations, PIPs move by at most
`r sprintf("%.3f", max(pip_df$range))` across three different seeds and
by at most `r sprintf("%.3f", max(cmp_prior$abs_diff))` between the
g-prior and the BIC-based prior. Given the number of MCMC iterations,
the search prove itself to be stable.

# 3. BAS vs Horseshoe: model comparison

Finally we compare the models selected by the two approaches, check on
which coefficient do they agree and on which they dont'.

```{r}
hs <- readRDS("../data/03_horseshoe_summary.rds")
hs_names_raw <- c("seat_confort", "time_convenience", "food_drink",
                   "gate_location", "wifi", "entertainment",
                   "online_support", "ease_booking", "onboard_service",
                   "legroom", "baggage", "checkin",
                   "cleanliness", "online_boarding", "Age_std", "FlightDist_std",
                   "Class_EcoPlus", "Class_Business", "logdep_std",
                   "TypeTravel_Personal", "CustType_disloyal", "logmid_delay")
kappa_ordered <- hs$post_kappa[c(paste0("beta[", 1:21, "]"), "beta_arr")]

names(kappa_ordered)
length(hs$post_kappa)

cmp_methods <- data.frame(term = hs_names_raw,
                           PIP = round(fit_glm_g$probne0[-1], 3),
                           kappa = round(kappa_ordered, 3))
cmp_methods$signal_BAS <- cmp_methods$PIP > 0.5
cmp_methods$signal_HS  <- cmp_methods$kappa < 0.5
cmp_methods$agree <- cmp_methods$signal_BAS == cmp_methods$signal_HS
agreement_rate <- mean(cmp_methods$agree)
print("two methods agreement rate is: ")
agreement_rate
```

```{r hs-comparison-plot, fig.width=7, fig.height=6}
cmp_methods %>%
  mutate(term = factor(term, levels = term[order(PIP)])) %>%
  ggplot(aes(x = PIP, y = 1 - kappa, color = agree)) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey60") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60") +
  scale_color_manual(values = c("TRUE" = "#3B6D11", "FALSE" = "#D85A30"),
                      labels = c("TRUE" = "Agree", "FALSE" = "Disagree"), name = NULL) +
  labs(title = "BAS posterior inclusion probability vs. horseshoe (1 - kappa)",
       subtitle = sprintf("%.0f%% agreement on signal (>0.5) vs. noise (<=0.5)", 100*agreement_rate),
       x = "PIP (BAS, g-prior)", y = expression(1 - kappa[j]~"(horseshoe)")) +
  theme_minimal(base_size = 10)
```

`r sprintf("%.0f%%", 100*agreement_rate)` of the 22 predictors get the
same signal/noise call from BAS's model-space search and from the
horseshoe's continuous shrinkage -- two structurally different
estimation strategies (discrete model averaging vs. a single shrinkage
prior) landing on the same handful of variables. The one disagreement,
`Class: Business` (PIP
`r cmp_methods$PIP[cmp_methods$term=="Class_Business"]`, $\kappa=$
`r cmp_methods$kappa[cmp_methods$term=="Class_Business"]`), is a genuine
borderline case in *both* methods (BAS just below its 0.5 cutoff, the
horseshoe's credible interval for it crossing zero in Section 03) -- not
a real contradiction between them.

## 4.Export results

```{r export-bas}
saveRDS(fit_glm_g, "../data/A01_BAS_model.rds")
```
