---
title: "0. Data audit, EDA, and the subsampling decision"
subtitle: "Airline Customer Satisfaction — Bayesian Learning project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## 1. Import and cleanup

We start, as usual, by clearing the workspace and fixing the seed for reproducibility.

```{r}
rm(list = ls())
set.seed(2026)

library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
```

The source file has 22 columns whose original names contain spaces and slashes
(`"Departure/Arrival time convenient"`, `"Departure Delay in Minutes"`, ...).
We rename them once, right away, to short `snake_case` names: every later
section (conjugate model, JAGS, BAS, ...) then refers to the same clean names
instead of re-deriving them each time.

```{r}
raw <- read.csv("Airline_customer_satisfaction.csv", stringsAsFactors = FALSE)

new_names <- c(
  "satisfaction", "customer_type", "age", "travel_type", "cabin_class",
  "flight_distance", "seat_comfort", "time_convenient", "food_drink",
  "gate_location", "wifi", "entertainment", "online_support", "ease_booking",
  "onboard_service", "legroom", "baggage", "checkin", "cleanliness",
  "online_boarding", "dep_delay", "arr_delay"
)
names(raw) <- new_names

dim(raw)
```

## 2. Data audit

A quick pass over structure, duplicates and missingness before anything else.

```{r}
str(raw)
cat("Duplicated rows:", sum(duplicated(raw)), "\n")

na_counts <- colSums(is.na(raw))
na_counts[na_counts > 0]
```

The only missingness in the whole dataset is in `arr_delay`
(`r na_counts["arr_delay"]` rows, `r round(100*na_counts["arr_delay"]/nrow(raw), 2)`%
of the data) — trivial to handle, we deal with it in Section 4.

The 14 service-quality items are Likert-type ratings; we check they all lie in
the advertised 0-5 range.

```{r}
ordinal_vars <- c("seat_comfort", "time_convenient", "food_drink", "gate_location",
                   "wifi", "entertainment", "online_support", "ease_booking",
                   "onboard_service", "legroom", "baggage", "checkin",
                   "cleanliness", "online_boarding")

sapply(raw[ordinal_vars], range)
```

All items range over 0-5 (`baggage` never actually hits 0). The remaining
numeric variables also look sane: `age` in
[`r min(raw$age)`, `r max(raw$age)`], `flight_distance` in
[`r min(raw$flight_distance)`, `r max(raw$flight_distance)`] miles, and both
delay variables non-negative.

We convert the four categorical variables to factors, fixing a natural order
for `cabin_class` (used again in Section 2, where we compare a nominal-dummy
vs. an ordinal-score encoding of class).

```{r}
raw$satisfaction  <- factor(raw$satisfaction, levels = c("dissatisfied", "satisfied"))
raw$customer_type <- factor(raw$customer_type)
raw$travel_type   <- factor(raw$travel_type)
raw$cabin_class   <- factor(raw$cabin_class, levels = c("Eco", "Eco Plus", "Business"))

round(prop.table(table(raw$satisfaction)), 3)
```

The response is close to balanced (`r round(100*mean(raw$satisfaction == "satisfied"),1)`%
satisfied), which is exactly the regime where a simple Bernoulli/logistic
likelihood is appropriate no need for anything fancier to handle imbalance.

We now drop the `r na_counts["arr_delay"]` rows with missing `arr_delay`
(0.3% of the data: dropping is simpler than imputing and changes nothing of
substance) and carry a `row_id` along, which will make the train/test split
in Section 5 trivial to construct.

```{r}
data <- raw %>% filter(!is.na(arr_delay))
data$row_id <- seq_len(nrow(data))
nrow(data)
```

## 3. Exploratory analysis

### 3.1 Correlation between the 14 service items

The 14 ordinal items are the core covariates for the horseshoe model in
Section 3 of the project, so it matters whether they are independent. We reuse
the correlation-heatmap recipe from the JAGS lab notes (`image()` +
`RColorBrewer`, rather than pulling in a new plotting package just for this).

```{r}
plot_corr_heatmap <- function(mat, labels = colnames(mat), title = "Correlation matrix") {
  pal <- colorRampPalette(rev(brewer.pal(11, "RdBu")))(100)
  layout(matrix(c(1, 2), nrow = 1), widths = c(4, 1))
  par(mar = c(7, 7, 4, 1))
  image(1:ncol(mat), 1:ncol(mat), mat,
        xlab = "", ylab = "", main = title,
        col = pal, zlim = c(-1, 1), axes = FALSE)
  axis(1, 1:ncol(mat), labels, las = 2, cex.axis = 0.8)
  axis(2, 1:ncol(mat), labels, las = 2, cex.axis = 0.8)
  box()
  # colorbar
  par(mar = c(7, 1, 4, 3))
  z <- seq(-1, 1, length.out = 100)
  image(1, z, matrix(z, nrow = 1), col = pal, axes = FALSE, xlab = "", ylab = "")
  axis(4, las = 2)
  box()
}

corr_mat <- cor(data[ordinal_vars])
plot_corr_heatmap(corr_mat, title = "Correlation between the 14 service-quality items")

# save a copy for the report
dir.create("figs", showWarnings = FALSE)
png("figs/correlation_heatmap.png", width = 8, height = 6, units = "in", res = 150)
plot_corr_heatmap(corr_mat, title = "Correlation between the 14 service-quality items")
invisible(dev.off())
```

Two clear blocks stand out, plus a third, weaker one:

-   **Digital/online cluster**: `wifi`, `online_support`, `ease_booking`,
    `online_boarding` (pairwise correlations
    `r round(min(corr_mat["wifi", c("online_support","ease_booking","online_boarding")]),2)`–`r round(max(corr_mat["wifi", c("online_support","ease_booking","online_boarding")]),2)`).
-   **Physical/onboard cluster**: `onboard_service`, `legroom`, `baggage`,
    `cleanliness` (pairwise correlations roughly 0.4–0.6).
-   **Boarding-logistics cluster**: `seat_comfort`, `time_convenient`,
    `food_drink`, `gate_location` also move together, most likely *because*
    they share the same block of structural zeros discussed in Section 4
    below (customers with an "N/A" on one tend to have it on all four).

None of this is surprising, but it is exactly the justification for the
horseshoe shrinkage prior later: several of these 14 covariates are
near-substitutes for one another, so a model that shrinks the redundant ones
towards zero is preferable to plain vague-prior logistic regression with all
14 included.

### 3.2 A few service items, by satisfaction

```{r, fig.width=7, fig.height=5}
box_df <- data %>%
  select(satisfaction, wifi, entertainment, onboard_service, cleanliness) %>%
  pivot_longer(-satisfaction, names_to = "item", values_to = "rating")

ggplot(box_df, aes(x = satisfaction, y = rating, fill = satisfaction)) +
  geom_boxplot(show.legend = FALSE) +
  facet_wrap(~ item) +
  labs(title = "Service ratings by satisfaction", x = NULL, y = "Rating (0-5)") +
  theme_minimal()
```

Satisfied customers rate every service item higher on average, as expected;
`wifi` and `entertainment` show the widest separation, hinting that they may
end up carrying real signal in the regression models later, rather than being
shrunk away.

## 4. The delay pair: decorrelating arrival from departure delay

```{r}
cor(data$dep_delay, data$arr_delay)
```

```{r, fig.width=6, fig.height=5}
ggplot(data, aes(x = dep_delay, y = arr_delay)) +
  geom_point(alpha = 0.03) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Arrival vs. departure delay",
       x = "Departure delay (min)", y = "Arrival delay (min)") +
  theme_minimal()
```

A correlation of `r round(cor(data$dep_delay, data$arr_delay), 3)` means
including both raw variables in a regression is close to duplicating one
predictor — exactly the near-collinearity the horseshoe prior is meant to
handle, but there is a cheaper fix for this *specific* pair. We regress
`arr_delay` on `dep_delay` and keep the **residual** as a second covariate:
by the defining property of least squares, a residual is orthogonal to every
column of the design matrix used to produce it, so this new variable is
uncorrelated with `dep_delay` *by construction* — no shrinkage prior needed to
discover that.

```{r}
delay_fit <- lm(arr_delay ~ dep_delay, data = data)
data$arr_delay_extra <- resid(delay_fit)

cor(data$arr_delay_extra, data$dep_delay)
```

As a sanity check, this is nearly the same information as the naive
`arr_delay - dep_delay` "minutes gained or lost in the air", since the
estimated slope is close to 1:

```{r}
coef(delay_fit)
excess <- data$arr_delay - data$dep_delay
cor(excess, data$arr_delay_extra)
```

We keep `arr_delay_extra` (exactly orthogonal to `dep_delay`) rather than the
naive difference (only approximately so) as the second delay covariate for
Sections 2-3, alongside `dep_delay` itself.

## 5. Structural zeros

```{r}
zero_share <- sort(sapply(data[ordinal_vars], function(x) mean(x == 0)), decreasing = TRUE)
round(zero_share, 4)
```

Four items — `time_convenient`, `food_drink`, `seat_comfort`, `entertainment`
have a non-trivial block of zeros (2-5% of rows), well above the near-zero
rates for the rest. A real "1" (very dissatisfied) and a coded "0"
(not applicable, e.g. no in-flight entertainment on that aircraft) are
different things, but disentangling them would need a zero-inflated / mixture
likelihood for each item. We flag this here and treat it as a **robustness
check** later (Section 5 of the project: refit excluding these rows and see
whether conclusions move), rather than building extra model complexity into
the main analysis now.

## 6. The subsampling decision

The full file has `r nrow(data)` rows (after dropping the 393 with missing
`arr_delay`), while the project brief describes a curated 1,000-row extract.
We draw a **stratified** training subsample of $n\approx 1200$ — matching the
brief's intended scale and keeping every MCMC run in later sections fast —
and keep essentially all remaining rows as a held-out test set for the
prediction exercise in Section 6. Stratifying on `satisfaction` and
`cabin_class` (6 strata) keeps the subsample representative of the population
on exactly the variables the whole analysis is about.

```{r}
stratified_sample <- function(pop, strata_vars, n_target, seed = 2026) {
  set.seed(seed)
  frac <- n_target / nrow(pop)
  pop %>%
    group_by(across(all_of(strata_vars))) %>%
    slice_sample(prop = frac) %>%
    ungroup()
}

train <- stratified_sample(data, c("satisfaction", "cabin_class"), n_target = 1200, seed = 2026)
test  <- data %>% filter(!row_id %in% train$row_id)

nrow(train); nrow(test)
```

A quick check that the sampling did what it should: stratum shares in `train`
should match the population shares closely.

```{r}
data   %>% count(satisfaction, cabin_class) %>% mutate(pop_share   = round(n / sum(n), 3)) %>% select(-n) %>%
  full_join(
    train %>% count(satisfaction, cabin_class) %>% mutate(train_share = round(n / sum(n), 3)) %>% select(-n),
    by = c("satisfaction", "cabin_class")
  )
```

They match to within a fraction of a percentage point in every stratum, as
they should by the law of large numbers applied within each group.

## 7. Save the split for later sections

```{r}
dir.create("data", showWarnings = FALSE)
write.csv(train %>% select(-row_id), "data/train_subsample.csv", row.names = FALSE)
write.csv(test  %>% select(-row_id), "data/test_holdout.csv",   row.names = FALSE)
```

From Section 1 onward we simply read `data/train_subsample.csv` (fit) and
`data/test_holdout.csv` (predict), so the subsampling never has to be redone.

## Summary

Before touching a single prior: the response is close to balanced, the 14
service ratings form 2-3 correlated blocks (motivating shrinkage rather than
using all 14 naively), the two delay variables are near-duplicates (fixed by
keeping `dep_delay` and an orthogonal residual `arr_delay_extra`), a handful
of items carry a structural-zero block worth revisiting later, and a stratified
~1200-row training subsample keeps the MCMC-based sections fast while
~`r nrow(test)` held-out rows give a low-variance out-of-sample check.
