---
title: "01 - Conjugate Baseline: How Satisfied Are Customers, Before Any Covariates?"
subtitle: "Airline Customer Satisfaction -- Bayesian Learning & Monte Carlo Simulations, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                       fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
```

This is the "Bayesian 101" opening beat: one exact, auditable number and its
uncertainty, with no black-box sampler yet. It uses the $n \approx 1{,}200$
stratified training subsample produced in `00_data_audit_and_eda.Rmd` --
closed-form throughout, since the Beta-Binomial model is conjugate.

```{r load-data}
train <- read.csv("airline_train_n1200.csv", stringsAsFactors = FALSE)
train$y <- as.integer(train$satisfaction == "satisfied")

n_train <- nrow(train)
```

# Model

$$
X \mid p \sim \text{Binomial}(n, p), \qquad p \sim \text{Beta}(\alpha_0, \beta_0),
$$

so that, by conjugacy,

$$
p \mid X \sim \text{Beta}(\alpha_0 + X,\ \beta_0 + n - X).
$$

We use $\text{Beta}(1,1)$ (uniform) as the default prior and revisit this
choice formally in the prior-sensitivity check below.

# Overall satisfaction rate

```{r overall-posterior}
a0 <- 1; b0 <- 1
X <- sum(train$y)
a_post <- a0 + X
b_post <- b0 + n_train - X

post_mean <- a_post / (a_post + b_post)
post_ci   <- qbeta(c(0.025, 0.975), a_post, b_post)
```

With `r n_train` customers in the training subsample and `r X` reporting
satisfaction, the posterior is
Beta(`r a_post`, `r b_post`), with posterior mean
`r sprintf("%.3f", post_mean)` and 95% credible interval
[`r sprintf("%.3f", post_ci[1])`, `r sprintf("%.3f", post_ci[2])`].

```{r overall-plot, fig.width=6, fig.height=4}
p_grid <- seq(0, 1, length.out = 1000)
plot_df <- data.frame(
  p = p_grid,
  prior     = dbeta(p_grid, a0, b0),
  posterior = dbeta(p_grid, a_post, b_post)
) %>% pivot_longer(-p, names_to = "distribution", values_to = "density")

ggplot(plot_df, aes(p, density, color = distribution)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c(prior = "grey60", posterior = "#185FA5")) +
  labs(title = "Overall satisfaction rate: prior vs. posterior",
       x = expression(p), y = "Density", color = NULL) +
  theme_minimal(base_size = 11)
```

# Subgroup posteriors

Each subgroup posterior is its own exact Beta -- no extra machinery is
needed, we simply condition on the rows in that subgroup.

```{r subgroup-helper}
subgroup_post <- function(data, var, a0 = 1, b0 = 1) {
  data %>%
    group_by(.data[[var]]) %>%
    summarise(n = n(), X = sum(y),
              a = a0 + X, b = b0 + n - X,
              mean = a / (a + b),
              lo95 = qbeta(0.025, a, b),
              hi95 = qbeta(0.975, a, b),
              .groups = "drop")
}

tbl_class    <- subgroup_post(train, "Class")
tbl_custtype <- subgroup_post(train, "Customer.Type")
tbl_travel   <- subgroup_post(train, "Type.of.Travel")
```

```{r subgroup-tables}
knitr::kable(tbl_class, digits = 3, caption = "Posterior satisfaction rate by Class")
knitr::kable(tbl_custtype, digits = 3, caption = "Posterior satisfaction rate by Customer Type")
knitr::kable(tbl_travel, digits = 3, caption = "Posterior satisfaction rate by Type of Travel")
```

```{r subgroup-plot, fig.width=9, fig.height=3.4}
make_density_df <- function(tbl, group_var) {
  tbl %>%
    rowwise() %>%
    reframe(group = .data[[group_var]],
            p = p_grid,
            density = dbeta(p_grid, a, b))
}

d_class    <- make_density_df(tbl_class, "Class") %>% mutate(facet = "Class")
d_custtype <- make_density_df(tbl_custtype, "Customer.Type") %>% mutate(facet = "Customer Type")
d_travel   <- make_density_df(tbl_travel, "Type.of.Travel") %>% mutate(facet = "Type of Travel")

bind_rows(d_class, d_custtype, d_travel) %>%
  ggplot(aes(p, density, color = group)) +
  geom_line(linewidth = 1) +
  facet_wrap(~ facet, scales = "free_y") +
  labs(title = "Subgroup posteriors for the satisfaction rate",
       x = expression(p), y = "Density", color = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom")
```

Business-class, loyal, and business-travel customers all show visibly
higher and non-overlapping posteriors than their counterparts -- exactly
the headline contrasts we quantify next.

# Monte Carlo subgroup contrasts

For two independent Beta posteriors we do not need a closed-form
comparison: we draw paired Monte Carlo samples from each and estimate
$P(p_A > p_B \mid \text{data})$ as the empirical proportion of draws where
one exceeds the other.

```{r mc-contrast}
set.seed(123)
S <- 200000

mc_contrast <- function(tbl, group_var, lvl1, lvl2, S = 200000) {
  r1 <- tbl[tbl[[group_var]] == lvl1, ]
  r2 <- tbl[tbl[[group_var]] == lvl2, ]
  d1 <- rbeta(S, r1$a, r1$b)
  d2 <- rbeta(S, r2$a, r2$b)
  list(draws1 = d1, draws2 = d2, prob = mean(d1 > d2))
}

contrast_class  <- mc_contrast(tbl_class, "Class", "Business", "Eco")
contrast_loyal  <- mc_contrast(tbl_custtype, "Customer.Type", "Loyal Customer", "disloyal Customer")
contrast_travel <- mc_contrast(tbl_travel, "Type.of.Travel", "Business travel", "Personal Travel")

contrast_summary <- data.frame(
  contrast = c("P(Business > Eco)", "P(Loyal > disloyal)", "P(Business travel > Personal travel)"),
  probability = c(contrast_class$prob, contrast_loyal$prob, contrast_travel$prob)
)
knitr::kable(contrast_summary, digits = 4)
```

```{r mc-diff-plot, fig.width=8, fig.height=3}
diff_df <- bind_rows(
  data.frame(diff = contrast_class$draws1 - contrast_class$draws2,  contrast = "Business - Eco"),
  data.frame(diff = contrast_loyal$draws1 - contrast_loyal$draws2,  contrast = "Loyal - disloyal"),
  data.frame(diff = contrast_travel$draws1 - contrast_travel$draws2, contrast = "Business travel - Personal travel")
)

ggplot(diff_df, aes(diff)) +
  geom_density(fill = "#3B6D11", alpha = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  facet_wrap(~ contrast, scales = "free") +
  labs(title = "Posterior distribution of the subgroup difference in satisfaction rate",
       x = expression(p[A] - p[B]), y = "Density") +
  theme_minimal(base_size = 9)
```

With essentially all posterior mass on the positive side for all three
contrasts, the data supports each direction with near certainty at this
sample size. This is still purely descriptive, though: the three grouping
variables are themselves correlated (e.g. business travellers are
disproportionately Business class), so we cannot yet say which of these
associations survive once we condition on the others -- that is exactly
the job of the regression in Section 02.

# Prior sensitivity analysis

We repeat the overall fit under three priors -- $\text{Beta}(1,1)$
(uniform), $\text{Beta}(0.5,0.5)$ (Jeffreys), and a mildly skeptical
$\text{Beta}(2,2)$ -- and also check the smallest subgroup we have
(`Eco Plus`, $n=$ `r tbl_class$n[tbl_class$Class == "Eco Plus"]`), where
prior sensitivity would show up first if it were going to show up at all.

```{r prior-sensitivity}
priors <- list(uniform = c(1, 1), jeffreys = c(0.5, 0.5), skeptical = c(2, 2))

sensitivity_row <- function(a0, b0, X, n) {
  a <- a0 + X; b <- b0 + n - X
  c(mean = a / (a + b), lo95 = qbeta(0.025, a, b), hi95 = qbeta(0.975, a, b))
}

sens_overall <- t(sapply(priors, function(pr) sensitivity_row(pr[1], pr[2], X, n_train)))

ecoplus <- train[train$Class == "Eco Plus", ]
n_ep <- nrow(ecoplus); X_ep <- sum(ecoplus$y)
sens_ecoplus <- t(sapply(priors, function(pr) sensitivity_row(pr[1], pr[2], X_ep, n_ep)))

knitr::kable(round(sens_overall, 4), caption = "Prior sensitivity -- overall satisfaction rate")
knitr::kable(round(sens_ecoplus, 4), caption = sprintf("Prior sensitivity -- Eco Plus subgroup (n=%d)", n_ep))
```

```{r sensitivity-plot, fig.width=8, fig.height=3.2}
sens_plot_df <- function(a0, b0, X, n, prior_name, group_name) {
  data.frame(p = p_grid, density = dbeta(p_grid, a0 + X, b0 + n - X),
             prior = prior_name, group = group_name)
}

sens_all <- bind_rows(
  lapply(names(priors), function(nm) sens_plot_df(priors[[nm]][1], priors[[nm]][2], X, n_train, nm, "Overall")),
  lapply(names(priors), function(nm) sens_plot_df(priors[[nm]][1], priors[[nm]][2], X_ep, n_ep, nm, "Eco Plus"))
)

ggplot(sens_all, aes(p, density, color = prior)) +
  geom_line(linewidth = 0.9) +
  facet_wrap(~ group, scales = "free") +
  labs(title = "Posterior under three priors: overall vs. smallest subgroup",
       x = expression(p), y = "Density", color = "Prior") +
  theme_minimal(base_size = 10)
```

Even for `Eco Plus`, the smallest and thus most prior-exposed subgroup, the
three posteriors are visually indistinguishable and the posterior means
differ only in the third decimal place. At this sample size the likelihood
already dominates any of these three reasonable priors -- a genuinely
uninformative result, in the sense that it tells us the *data*, not our
prior choice, is driving the conclusions.
