---
title: "1. Bernoulli/logit regression: reference prior and Zellner's g-prior"
subtitle: "Airline Customer Satisfaction — Bayesian Learning project"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 4.5)
```

The response `satisfaction` is binary, so we use a GLM: a **Bernoulli
likelihood with a logit link**,

$$\begin{align*}
  y_i \mid p_i &\overset{ind}\sim \mathscr{B}e(p_i), &\qquad i = 1,\dots,n \\
  \log\!\left(\frac{p_i}{1-p_i}\right) &= \beta_0 + x_i^T\beta .
\end{align*}$$

Taking this models as reference, we tune the first two simple models
(model selection methods will follow).

```{r libs}
rm(list = ls())
set.seed(2026)

library(rjags)
library(coda)
library(ggplot2)
library(dplyr)
library(tidyr)
```

## 0. Data transformation

```{r load}
train <- read.csv("data/train_subsample.csv", stringsAsFactors = FALSE)
test  <- read.csv("data/test_holdout.csv",    stringsAsFactors = FALSE)
```

The response is coded as $1 =$ `satisfied`, $0 =$ `dissatisfied`.

```{r response}
to_y <- function(df) {
  as.integer(factor(df$satisfaction, levels = c("dissatisfied", "satisfied"))) - 1
}
train$y <- to_y(train)
test$y  <- to_y(test)

# double-check to assess that response class balance is preserved
round(c(train = mean(train$y), test = mean(test$y)), 4)
```

### One-hot encoding of categorical features

The three categorical variables are `customer_type` (2 levels),
`travel_type` (2 levels) and `cabin_class` (3 levels). Apply one-hot
encoding to translate the classes in dummies to account correctly their
relevance in the prediction.

To avoid ending up with a feature matrix non-invertible, we treat each
category, which now is a single dummy with the relative coefficient, as
a log-odd contrast again a reference category. See below the reference
used.

| variable        | reference level   | dummies created                   |
|-----------------|-------------------|-----------------------------------|
| `customer_type` | `Loyal Customer`  | `cust_disloyal`                   |
| `travel_type`   | `Business travel` | `travel_personal`                 |
| `cabin_class`   | `Eco`             | `class_ecoplus`, `class_business` |

```{r encode}
num_vars <- c("age", "flight_distance",
              "seat_comfort", "time_convenient", "food_drink", "gate_location",
              "wifi", "entertainment", "online_support", "ease_booking",
              "onboard_service", "legroom", "baggage", "checkin",
              "cleanliness", "online_boarding",
              "dep_delay", "arr_delay_extra")

# function to encode categorical variables
encode <- function(df) {
  D <- data.frame(
    cust_disloyal   = as.integer(df$customer_type == "disloyal Customer"),
    travel_personal = as.integer(df$travel_type   == "Personal Travel"),
    class_ecoplus   = as.integer(df$cabin_class   == "Eco Plus"),
    class_business  = as.integer(df$cabin_class   == "Business")
  )
  cbind(D, df[, num_vars]) # with this 
}

Xtr <- encode(train)
Xte <- encode(test)

# shape new datasets, ready to be used in prediction
dtr <- cbind(y = train$y, Xtr)
dte <- cbind(y = test$y,  Xte)

str(dtr[, 1:6])
```

Note how keeping both `dep_delay` and `arr_delay` would make $X^TX$
almost singular, with huge problems for the Zellner's prior due the
necessity of inverting the matrix.

## 1. Diagnostic tools

A bunch of functions as helper for the prediction model's diagnostics.

In L06 the diagnostic was "residuals vs. $x$ with a loess curve". For a
Bernoulli response that plot is useless: $y_i - \hat p_i$ can only take
the two values $\hat p_i$ and $\hat p_i - 1$, so the scatter always
splits into two curves regardless of how good the model is. The standard
replacement is the **binned residual plot**: sort the observations by
fitted probability, group them into bins, and plot the *average*
residual per bin. Under a correct model the binned residuals are centred
at zero and roughly $95\%$ of them fall inside $\pm 2\,\mathrm{SE}$,
with $\mathrm{SE} = \sqrt{\overline{p(1-p)}/m}$ for a bin of size $m$.

```{r diag-functions}
# --- discrimination and calibration metrics -------------------------------
auc_fun <- function(y, p) {                 # Mann-Whitney U form of the AUC
  r  <- rank(p)
  n1 <- as.numeric(sum(y == 1)); n0 <- as.numeric(sum(y == 0))
  (sum(r[y == 1]) - n1 * (n1 + 1) / 2) / (n1 * n0)
}

metrics <- function(y, p, thr = 0.5) {
  yhat <- as.integer(p > thr)
  data.frame(
    accuracy = mean(yhat == y),
    auc      = auc_fun(y, p),
    brier    = mean((p - y)^2),
    logloss  = -mean(y * log(pmax(p, 1e-12)) + (1 - y) * log(pmax(1 - p, 1e-12))),
    sens     = mean(yhat[y == 1] == 1),
    spec     = mean(yhat[y == 0] == 0)
  )
}

confusion <- function(y, p, thr = 0.5) {
  table(predicted = factor(as.integer(p > thr), levels = c(0, 1),
                           labels = c("dissatisfied", "satisfied")),
        observed  = factor(y, levels = c(0, 1),
                           labels = c("dissatisfied", "satisfied")))
}

# --- curves ---------------------------------------------------------------
roc_df <- function(y, p, n_grid = 500) {     # thresholds on a quantile grid
  thr <- unique(quantile(p, probs = seq(0, 1, length.out = n_grid), names = FALSE))
  data.frame(
    thr = thr,
    tpr = sapply(thr, function(t) mean(p[y == 1] > t)),
    fpr = sapply(thr, function(t) mean(p[y == 0] > t))
  )
}

calib_df <- function(y, p, bins = 20) {      # observed frequency vs predicted p
  b <- cut(p, breaks = quantile(p, probs = seq(0, 1, length.out = bins + 1)),
           include.lowest = TRUE)
  data.frame(p = p, y = y, bin = b) %>%
    group_by(bin) %>%
    summarise(p_hat = mean(p), freq = mean(y), m = n(), .groups = "drop") %>%
    mutate(se = sqrt(pmax(freq * (1 - freq), 1e-12) / m))
}

binned_resid_df <- function(y, p, bins = 40) {
  b <- cut(p, breaks = quantile(p, probs = seq(0, 1, length.out = bins + 1)),
           include.lowest = TRUE)
  data.frame(p = p, r = y - p, bin = b) %>%
    group_by(bin) %>%
    summarise(p_hat = mean(p), resid = mean(r), m = n(),
              se = 2 * sqrt(mean(p * (1 - p)) / n()), .groups = "drop")
}

# --- plots ----------------------------------------------------------------
plot_roc <- function(d, title) {
  ggplot(d, aes(fpr, tpr)) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", colour = "grey50") +
    geom_line(colour = "steelblue", linewidth = 0.9) +
    coord_equal() + labs(title = title, x = "False positive rate",
                         y = "True positive rate") + theme_minimal()
}

plot_calib <- function(d, title) {
  ggplot(d, aes(p_hat, freq)) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", colour = "grey50") +
    geom_errorbar(aes(ymin = freq - 2 * se, ymax = freq + 2 * se),
                  width = 0, colour = "grey60") +
    geom_point(colour = "steelblue", size = 2) +
    labs(title = title, x = "Predicted probability (bin mean)",
         y = "Observed frequency") + theme_minimal()
}

plot_binned <- function(d, title) {
  ggplot(d, aes(p_hat, resid)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "red") +
    geom_point(size = 1.8, colour = "steelblue") +
    labs(title = title, x = "Fitted probability (bin mean)",
         y = "Mean residual") + theme_minimal()
}

# --- coefficient table ----------------------------------------------------
coef_table <- function(mean, sd, names, level = 0.95) {
  z <- qnorm(1 - (1 - level) / 2)
  data.frame(term = names, post_mean = mean, post_sd = sd,
             lower = mean - z * sd, upper = mean + z * sd,
             row.names = NULL)
}

# plot credibile intervals of the estimated coefficients
plot_coefs <- function(tab, title) {
  tab %>% filter(term != "Intercept") %>%
    mutate(term = factor(term, levels = rev(term))) %>%
    ggplot(aes(x = post_mean, y = term)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
    geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0, colour = "grey50") +
    geom_point(colour = "steelblue", size = 2) +
    labs(title = title, x = expression(beta[j]), y = NULL) + theme_minimal()
}

```

## 2. MCMC setup

Since there are features living in different scales we standardize them,
therefore boosting the convergence of the method of the sampler and
reducing autocorrelation (this won't affect the priors since both the
uninformative and the Zellner's are scale-invariant).

```{r}
Z_train  <- scale(as.matrix(Xtr))
x_center <- attr(Z_train, "scaled:center")
x_scale  <- attr(Z_train, "scaled:scale")

# the test set must use the TRAINING mean and sd, never its own
Z_test <- scale(as.matrix(Xte), center = x_center, scale = x_scale)

n <- nrow(Z_train)   # 1197 observations
k <- ncol(Z_train)   # 22 covariates

# MCMC settings
n.chains <- 3
n.adapt  <- 1000
burn     <- 1000
n.iter   <- 4000
thin     <- 2

########## MCMC helper functions ############################################

# --- fit a model using JAGS ---
fit_jags <- function(bugs_code, extra_data = list()) {
  
  # reproducible chains: fix the RNG of each chain explicitly
  inits <- lapply(seq_len(n.chains), function(c)
    list(.RNG.name = "base::Mersenne-Twister", .RNG.seed = 2026 + c))
  
  model <- jags.model(textConnection(bugs_code),
                      data     = c(list(y = dtr$y, Z = Z_train, n = n, k = k),
                                   extra_data),
                      inits    = inits,
                      n.chains = n.chains, n.adapt = n.adapt, quiet = TRUE)
  update(model, n.iter = burn)                      # burn-in
  coda.samples(model, variable.names = c("beta0", "beta"),
               n.iter = n.iter, thin = thin)
}

# --- convergence  JAGS summary ---
mcmc_conv_diag <- function(samples) {
  data.frame(min_ESS  = round(min(effectiveSize(samples))),
             max_Rhat = round(max(gelman.diag(samples,
                                              multivariate = FALSE)$psrf[, 1]), 4))
}

# --- back-transform the draws to the original covariate scale --------------
to_original <- function(samples) {
  S  <- as.matrix(samples)
  b0 <- S[, "beta0"]
  B  <- S[, paste0("beta[", 1:k, "]"), drop = FALSE]
  
  B_orig  <- sweep(B, 2, x_scale, "/")             # beta_j / s_j
  b0_orig <- b0 - as.vector(B_orig %*% x_center)   # intercept correction
  
  out <- cbind(b0_orig, B_orig)
  colnames(out) <- c("Intercept", names(Xtr))
  out
}

# --- posterior summary in the format expected by plot_coefs() --------------
post_table <- function(samples) {
  O <- to_original(samples)
  data.frame(term      = colnames(O),
             post_mean = colMeans(O),
             post_sd   = apply(O, 2, sd),
             lower     = apply(O, 2, quantile, 0.025),
             upper     = apply(O, 2, quantile, 0.975),
             row.names = NULL)
}

# --- posterior predictive probability E[p_new | y] -------------------------
# We average the probabilities over the posterior draws, which is the Bayesian
# point prediction; the test set never enters JAGS, we only push the draws
# through the link in R. 500 draws are plenty for a posterior *mean*.
post_pred <- function(samples, Znew, n_use = 500) {
  S   <- as.matrix(samples)
  b0  <- S[, "beta0"]
  B   <- S[, paste0("beta[", 1:k, "]"), drop = FALSE]
  idx <- round(seq(1, length(b0), length.out = min(n_use, length(b0))))
  
  acc <- numeric(nrow(Znew))
  for (s in idx) acc <- acc + plogis(b0[s] + as.vector(Znew %*% B[s, ]))
  return (acc / length(idx))
}
```

## 3. Non-informative prior model

We put a flat prior on the whole coefficient vector,

$$\pi(\beta_0, \beta) \propto
1,$$

The posterior is then simply proportional to the likelihood,

$$\pi(\beta \mid y, X) \;\propto\;
\prod_{i=1}^n p_i^{y_i}(1-p_i)^{1-y_i},
\qquad p_i = \frac{1}{1+e^{-x_i^T\beta}} .$$

However, JAGS doesn't have a "proper" improper prior, so we have to rely
on a very flat normal gaussian, with a sd that makes the prior de facto
flat in the region of interest.

```{r}
model_flat <- "model{

  # Likelihood: Bernoulli with logit link
  for(i in 1:n){
    y[i] ~ dbern(p[i])
    logit(p[i]) <- beta0 + inprod(Z[i,], beta[])
  }

  # Non-informative prior: numerically flat (nano-like scale on deviation)
  beta0 ~ dnorm(0, 1.0E-9)
  for(j in 1:k){
    beta[j] ~ dnorm(0, 1.0E-9)
  }
}"

samples_flat <- fit_jags(model_flat)
```

### non-informative MCMC diagnostic

```{r, fig.height=14, fig.width=8}
plot(samples_flat)
```

```{r, fig.height=14, fig.width=8}
autocorr.plot(samples_flat, ask = FALSE)
```

```{r}
mcmc_conv_diag(samples_flat)
```

### non-informative model summary

```{r}
# coefficient's table
tab_flat <- post_table(samples_flat)
knitr::kable(tab_flat, digits = 4,
             caption = "Reference prior: posterior mean, sd and 95% credible intervals")

# forest plot
plot_coefs(tab_flat, "Reference prior: posterior means and 95% credible intervals")
```

```{r, fig.height=14, fig.width=8}
# here the desnity is on the original scale, and it's a little better-looking than the base plot of R

draws_flat <- to_original(samples_flat)

#show <- c("cust_disloyal", "class_business", "wifi", "entertainment") 
as.data.frame(draws_flat) %>%   
  pivot_longer(everything(), names_to = "term", values_to = "beta") %>% 
  ggplot(aes(beta)) +   
  geom_density(fill = "steelblue", alpha = 0.4) +   
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +   
  facet_wrap(~ term, scales = "free") +   
  labs(title = "Reference prior: posterior densities", x = expression(beta[j])) + 
  theme_minimal()
```

```{r}
or <- exp(draws_flat)
data.frame(term = colnames(draws_flat),
           OR_mean = colMeans(or),
           t(apply(or, 2, quantile, c(0.025, 0.975)))) %>%
  knitr::kable(digits = 3, row.names = FALSE,
               caption = "Posterior of the odds ratios exp(beta_j)")
```

### non-informative diagnostic

```{r}
p_tr_flat <- post_pred(samples_flat, Z_train)
p_te_flat <- post_pred(samples_flat, Z_test)

ev_flat <- rbind(
  cbind(set = "train", metrics(dtr$y, p_tr_flat)),
  cbind(set = "test",  metrics(dte$y, p_te_flat))
)
knitr::kable(ev_flat, digits = 4, caption = "Reference prior: performance")
```

```{r}
plot_binned(binned_resid_df(dtr$y, p_tr_flat),
            "Binned residuals, training set (reference prior)")
```

```{r}
confusion(dte$y, p_te_flat)

gridExtra::grid.arrange(
  plot_roc(roc_df(dte$y, p_te_flat), "ROC, held-out test set"),
  plot_calib(calib_df(dte$y, p_te_flat), "Calibration, held-out test set"),
  ncol = 2)
```

## 3. Zellner's g-prior model

For a Gaussian likelihood the scale factor is
\$\\sigma^2\$.\ A\ Bernoulli\ likelihood\ has\ no\ \$\\sigma^2\$, so the
standard generalisation replaces $X^TX$ by the **Fisher information
evaluated under the null model**: with only an intercept,
\$\\hat{p}\_i=\\bar{y}\$ and the information is
\$\\bar{w}\\tilde{Z}\^T\\tilde{Z}\$ with
\$\\bar{w}=\\bar{y}(1-\\bar{y})\$. The prior is therefore

$$\beta\ |\ \alpha \sim \mathcal{N}_k\left( 0, \frac{\alpha}{\bar{w}}(\tilde{Z}^T\tilde{Z}) \right), \quad \pi(\beta_0) \propto 1$$

that is, a *precision* matrix
\$\\Lambda=\\frac{\\bar{w}}{\\alpha}\\tilde{Z}\^T\\tilde{Z}\$. The
precision formulatio is convinient since JAGS use it, so we can avoid
matrix inversion.

The interpretation of
$\alpha$ carries over: the prior carries $\frac{1}{\alpha}$ times the
information of the whole sample, i.e. it is worth $\frac{n}{\alpha}$
observations. Hence \$\\alpha=n\$ is the **unit information** prior (the
prior weighs as much as a single observation), \$\\alpha=1\$ gives prior
and sample equal weight, and \$\\alpha\\to\\infty\$ recovers Model 1.

```{r}
ZtZ  <- crossprod(Z_train)                       # t(Z) %*% Z
wbar <- mean(dtr$y) * (1 - mean(dtr$y))          # null-model weight

model_zellner <- "model{

  # Likelihood: identical to Model 1
  for(i in 1:n){
    y[i] ~ dbern(p[i])
    logit(p[i]) <- beta0 + inprod(Z[i,], beta[])
  }

  # Zellner's g-prior: multivariate normal with precision Lambda, passed as data
  beta[1:k] ~ dmnorm(zeros[], Lambda[,])

  # Flat prior on the intercept
  beta0 ~ dnorm(0, 1.0E-6)
}"

alphas <- c(0.01, 0.1, 1, 10, n)

fits_z <- lapply(alphas, function(a)
  fit_jags(model_zellner,
           extra_data = list(zeros  = rep(0, k),
                             Lambda = (wbar / a) * ZtZ)))
names(fits_z) <- alphas
```

### Zellner's MCMC convergence

```{r}
diag_z <- do.call(rbind, Map(function(s, a) cbind(alpha = a, mcmc_conv_diag(s)),
                             fits_z, alphas))
knitr::kable(diag_z, row.names = FALSE,
             caption = "Convergence across the alpha grid")
```

```{r, fig.height=14, fig.width=8}
# just as a check for the first model (maybe you can improve this)
plot(fits_z[["0.01"]])
```

```{r, fig.height=14, fig.width=8}
autocorr.plot(fits_z[["0.01"]], ask = FALSE)
```

in case we're able to get better mixing, continue with this prior
