---
title: "02 - Core Bayesian Logistic Regression (JAGS, Vague Priors)"
subtitle: "Airline Customer Satisfaction -- Bayesian Learning & Monte Carlo Simulations, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                       fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(rjags)
library(coda)
library(loo)
```

This is the model the audience can hold in their head: a handful of
interpretable odds ratios. It is the interpretable fallback that every
more elaborate model in later sections has to outperform to earn its
place. We follow the Bernoulli-regression / Titanic-survival template
from the JAGS notes (binary outcome, mix of categorical and continuous
predictors) almost line for line.

```{r load-data}
train <- read.csv("airline_train_n1200.csv", stringsAsFactors = FALSE)
train$y <- as.integer(train$satisfaction == "satisfied")
n_train <- nrow(train)
```

# Covariate preparation

Covariates: `Age`, `Seat comfort`, `Flight Distance`, `Class`, the two
delay variables from Section 00 (`Departure Delay` and the
decorrelated `ArrivalDelay.Residual`), `Type of Travel`, `Customer Type`.

The original plan was to log-transform `Flight Distance` together with
the delay variables. Before doing that blindly, we check the actual skew
on the training subsample:

```{r skew-check}
skew <- function(x) { x <- x[!is.na(x)]; m <- mean(x); s <- sd(x); mean(((x - m) / s)^3) }

signedLogArr <- sign(train$ArrivalDelay.Residual) * log1p(abs(train$ArrivalDelay.Residual))

skew_tbl <- data.frame(
  variable = c("Flight Distance", "log1p(Flight Distance)",
               "Departure Delay", "log1p(Departure Delay)",
               "Arrival Delay residual", "signed-log(Arrival Delay residual)"),
  skewness = round(c(skew(train$Flight.Distance), skew(log1p(train$Flight.Distance)),
                      skew(train$Departure.Delay.in.Minutes), skew(log1p(train$Departure.Delay.in.Minutes)),
                      skew(train$ArrivalDelay.Residual), skew(signedLogArr)), 3)
)
knitr::kable(skew_tbl)
```

`Flight Distance` is only mildly skewed (0.36) and a log transform
*overcorrects* it into a left skew (-1.70) -- so, contrary to the original
plan, we keep it on the raw scale. The two delay variables are genuinely
skewed (4.28 and 1.80) and both transforms clearly help: `log(1+x)` for
`Departure Delay` (non-negative), and a **signed log**,
$\text{sign}(x)\log(1+|x|)$, for the delay residual, which can be negative.
This mirrors the standard trick for a skewed variable that isn't bounded
below by zero.

```{r covariate-prep}
std <- function(x) as.numeric(scale(x))

logDepDelay     <- log1p(train$Departure.Delay.in.Minutes)
Age_std         <- std(train$Age)
Seat_std        <- std(train$Seat.comfort)
FlightDist_std  <- std(train$Flight.Distance)
logDepDelay_std <- std(logDepDelay)

# Arrival delay residual: standardize using the observed rows only, keep the
# single NA -- it is handled inside the JAGS model below, not dropped here.
obs  <- !is.na(signedLogArr)
mu_a <- mean(signedLogArr[obs]); sd_a <- sd(signedLogArr[obs])
arr_delay_std <- (signedLogArr - mu_a) / sd_a

n_missing_arr <- sum(is.na(arr_delay_std))

Class_EcoPlus   <- as.integer(train$Class == "Eco Plus")
Class_Business  <- as.integer(train$Class == "Business")
Class_score_std <- std(as.integer(factor(train$Class, levels = c("Eco", "Eco Plus", "Business"), ordered = TRUE)))

TypeTravel_Personal <- as.integer(train$Type.of.Travel == "Personal Travel")
CustType_disloyal   <- as.integer(train$Customer.Type == "disloyal Customer")

X_dummy <- cbind(Age_std, Seat_std, FlightDist_std, Class_EcoPlus, Class_Business,
                  logDepDelay_std, TypeTravel_Personal, CustType_disloyal)
X_ordinal <- cbind(Age_std, Seat_std, FlightDist_std, Class_score_std,
                    logDepDelay_std, TypeTravel_Personal, CustType_disloyal)
```

`Class` is fit two ways for comparison below: two dummies
(`Eco Plus`, `Business`, reference = `Eco`) versus a single centred
ordinal score (1/2/3). Reference levels for the binary covariates are
the majority group in each case: `Business travel` and `Loyal Customer`.
The delay residual has `r n_missing_arr` missing value(s) in this
subsample; it is passed to JAGS as a partly-missing vector with its own
sampling statement, so the single NA is imputed as a genuine posterior
draw rather than silently dropped or mean-filled.

# Model specification

$$
y_i \mid p_i \sim \text{Bernoulli}(p_i), \qquad
\text{logit}(p_i) = \beta_0 + x_i\beta + \beta_{\text{arr}}\, \text{arr\_delay}_i,
\qquad \beta_0, \beta_j, \beta_{\text{arr}} \sim \mathscr{N}(0, 100).
$$

The arrival-delay residual is kept outside the main design matrix and
given its own coefficient together with an explicit sampling statement
`arr_delay[i] ~ dnorm(0, 1)` for **every** row -- for observed rows this
just re-states the (already standardized, so unit-variance) marginal
distribution; for the one missing row, this is exactly what lets JAGS
draw it as a latent parameter, informed by both this marginal
distribution and its role in the logistic regression likelihood, rather
than requiring the row to be dropped.

```{r model-code}
model_code <- function(link = "logit") {
  link_line <- if (link == "logit") "logit(p[i]) <- lin[i]" else "probit(p[i]) <- lin[i]"
  sprintf('model {
  for (i in 1:n) {
    y[i] ~ dbern(p[i])
    %s
    lin[i] <- beta0 + inprod(X[i,], beta) + beta_arr * arr_delay[i]
    log_lik[i] <- logdensity.bern(y[i], p[i])
    arr_delay[i] ~ dnorm(0, 1)
  }
  beta0 ~ dnorm(0, 0.01)
  beta_arr ~ dnorm(0, 0.01)
  for (j in 1:P) { beta[j] ~ dnorm(0, 0.01) }
}', link_line)
}
```

`log_lik[i] <- logdensity.bern(y[i], p[i])` is monitored throughout so
that WAIC can be computed via the `loo` package for every model fit
below, on a common footing.

```{r fit-function}
fit_model <- function(X, arr_delay, link = "logit",
                       n.chains = 3, n.adapt = 500, burn = 1500, n.iter = 3000, thin = 2) {
  dataList <- list(y = train$y, X = X, arr_delay = arr_delay, n = nrow(X), P = ncol(X))
  m <- jags.model(textConnection(model_code(link)), data = dataList,
                   n.chains = n.chains, n.adapt = n.adapt, quiet = TRUE)
  update(m, burn)
  coda.samples(m, variable.names = c("beta0", "beta", "beta_arr", "log_lik"),
               n.iter = n.iter, thin = thin)
}

extract_loglik <- function(samples) {
  sm <- as.matrix(samples)
  sm[, grepl("^log_lik", colnames(sm))]
}
```

# Choosing the `Class` encoding

```{r fit-encoding, cache=TRUE}
set.seed(123)
fit_dummy_logit   <- fit_model(X_dummy,   arr_delay_std, "logit")
fit_ordinal_logit <- fit_model(X_ordinal, arr_delay_std, "logit")

w_dummy   <- waic(extract_loglik(fit_dummy_logit))
w_ordinal <- waic(extract_loglik(fit_ordinal_logit))
cmp_encoding <- loo_compare(list(dummy = w_dummy, ordinal = w_ordinal))
```

```{r encoding-waic-print}
cat("Dummy encoding:\n");   print(w_dummy)
cat("\nOrdinal encoding:\n"); print(w_ordinal)
cat("\nComparison (positive elpd_diff favors the row):\n"); print(cmp_encoding)
```

The dummy encoding has the lower WAIC
(`r sprintf("%.1f", w_dummy$estimates["waic","Estimate"])` vs
`r sprintf("%.1f", w_ordinal$estimates["waic","Estimate"])`), though by
less than 2 standard errors of the difference -- a real but modest
preference, not an overwhelming one. Given `Class` is nominal rather than
genuinely equally-spaced ordinal, and the dummy encoding is both the more
flexible specification and the one that will let us report a separate,
directly interpretable odds ratio per class, we adopt it going forward.

# Choosing the link function

```{r fit-link, cache=TRUE}
set.seed(124)
fit_dummy_probit <- fit_model(X_dummy, arr_delay_std, "probit")
w_probit <- waic(extract_loglik(fit_dummy_probit))
cmp_link <- loo_compare(list(logit = w_dummy, probit = w_probit))
```

```{r link-waic-print}
cat("Probit link (dummy encoding):\n"); print(w_probit)
cat("\nComparison:\n"); print(cmp_link)
```

Logit is (very slightly) preferred, as textbook theory would predict for
this kind of data -- logit and probit links are nearly indistinguishable
in fit and only differ in the tail behaviour of the link function. We
carry the **logit link with dummy `Class` encoding** forward as the main
model for the rest of the project (this is also the specification the
horseshoe model in Section 03 builds on).

# Diagnostics for the final model

```{r diagnostics}
final_fit <- fit_dummy_logit
beta_cols <- colnames(as.matrix(final_fit))[!grepl("^log_lik", colnames(as.matrix(final_fit)))]

gd  <- gelman.diag(final_fit[, beta_cols], multivariate = FALSE)
ess <- effectiveSize(final_fit[, beta_cols])

cat("Max Gelman-Rubin R-hat across beta parameters:", round(max(gd$psrf[, 1]), 4), "\n")
cat("\nEffective sample sizes:\n"); print(round(ess))
```

```{r trace-plots, fig.width=8, fig.height=5}
par(mfrow = c(2, 2))
plot(as.mcmc.list(final_fit)[, "beta0"], trace = TRUE, density = FALSE, main = "beta0 (intercept)", auto.layout = FALSE)
plot(as.mcmc.list(final_fit)[, "beta[5]"], trace = TRUE, density = FALSE, main = "beta[5] (Class: Business)", auto.layout = FALSE)
acf(as.matrix(final_fit[[1]])[, "beta0"], main = "ACF: beta0, chain 1")
acf(as.matrix(final_fit[[1]])[, "beta[5]"], main = "ACF: beta[5], chain 1")
par(mfrow = c(1, 1))
```

Max R-hat is `r sprintf("%.3f", max(gd$psrf[, 1]))`, well below the usual
1.05 threshold, and effective sample sizes range from
`r round(min(ess))` (the intercept, as usual the slowest-mixing parameter
since it isn't itself standardized like the continuous covariates) to
`r round(max(ess))`. No further reparametrisation (e.g. centering the
0/1 dummies) is needed at this sample size.

# Posterior inference: odds ratios

```{r odds-ratios}
sm <- as.matrix(final_fit)
sm_beta <- sm[, beta_cols]

term_labels <- c(
  "beta0" = "Intercept",
  "beta[1]" = "Age",
  "beta[2]" = "Seat comfort",
  "beta[3]" = "Flight Distance",
  "beta[4]" = "Class: Eco Plus (vs Eco)",
  "beta[5]" = "Class: Business (vs Eco)",
  "beta[6]" = "log(1+Departure Delay)",
  "beta[7]" = "Personal Travel (vs Business travel)",
  "beta[8]" = "disloyal Customer (vs Loyal)",
  "beta_arr" = "Arrival delay residual (signed-log)"
)

post_mean <- colMeans(sm_beta)
post_ci   <- apply(sm_beta, 2, quantile, c(0.025, 0.975))

or_tbl <- data.frame(
  term = term_labels[names(post_mean)],
  beta_mean = round(post_mean, 3),
  beta_lo = round(post_ci[1, ], 3), beta_hi = round(post_ci[2, ], 3),
  OR = round(exp(post_mean), 3),
  OR_lo = round(exp(post_ci[1, ]), 3), OR_hi = round(exp(post_ci[2, ]), 3)
) %>% arrange(desc(OR))

knitr::kable(or_tbl, row.names = FALSE)
```

```{r or-plot, fig.width=7, fig.height=4.5}
or_tbl %>%
  filter(term != "Intercept") %>%
  mutate(term = factor(term, levels = term[order(OR)]),
         significant = OR_lo > 1 | OR_hi < 1) %>%
  ggplot(aes(x = OR, y = term, color = significant)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = OR_lo, xmax = OR_hi), height = 0.2) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c("TRUE" = "#185FA5", "FALSE" = "grey60"), guide = "none") +
  scale_x_log10() +
  labs(title = "Posterior odds ratios, logit model with dummy Class encoding",
       subtitle = "95% credible intervals; dashed line = no effect",
       x = "Odds ratio (log scale)", y = NULL) +
  theme_minimal(base_size = 11)
```

# Interpretation

`Business` class carries by far the largest effect
(OR $\approx$ `r or_tbl$OR[or_tbl$term == "Class: Business (vs Eco)"]`,
95% CI
[`r or_tbl$OR_lo[or_tbl$term == "Class: Business (vs Eco)"]`,
`r or_tbl$OR_hi[or_tbl$term == "Class: Business (vs Eco)"]`]),
confirming the Section 01 subgroup contrast even after adjusting for
everything else. `disloyal Customer` shows a strong *negative* effect
(OR $\approx$ `r or_tbl$OR[or_tbl$term == "disloyal Customer (vs Loyal)"]`),
also consistent with Section 01. `Seat comfort` is the strongest
continuous driver.

The interesting reversal is `Type of Travel`: Section 01 found an
essentially certain marginal association (personal-travel customers
looked less satisfied), but here, once `Class` and the other covariates
are held fixed, its 95% credible interval for `Personal Travel`
(`r or_tbl$OR_lo[or_tbl$term == "Personal Travel (vs Business travel)"]`
to
`r or_tbl$OR_hi[or_tbl$term == "Personal Travel (vs Business travel)"]`)
comes close to, but does not quite exclude, 1. This is exactly the
confounding flagged at the end of Section 01: business travellers are
disproportionately booked in Business class, and once class is in the
model directly, most of the apparent travel-type effect is explained
away. `Flight Distance` and `Departure Delay` point in an intuitive
direction but their credible intervals are wide enough that we would not
call either one a confident driver on its own -- Section 03's shrinkage
model, with the full battery of 14 service ratings added, will help
clarify which of these borderline effects survive alongside the stronger
correlated service signals.

# Export for later sections

Saved for reuse in Section 05 (WAIC/PSIS-LOO comparison and the
prior-variance sweep): the compact `loo`/`waic` result objects. Also
saved here for Section 06: a thinned set of posterior draws of `beta`,
and -- critically -- the exact training-set mean/SD used to standardize
every continuous covariate, so that test-set predictions apply the
*identical* transform rather than leaking test-set statistics into
preprocessing.

```{r export-loo}
final_loglik <- extract_loglik(final_fit)
loo_final <- loo::loo(final_loglik)
waic_final <- loo::waic(final_loglik)

# Thin to at most 1000 draws for compactness; plenty for posterior-mean
# predictions and calibration in Section 06.
sm_final <- as.matrix(final_fit)
keep_idx <- round(seq(1, nrow(sm_final), length.out = min(1000, nrow(sm_final))))
beta_draws <- sm_final[keep_idx, beta_cols]

standardization <- list(
  Age_mean = mean(train$Age), Age_sd = sd(train$Age),
  Seat_mean = mean(train$Seat.comfort), Seat_sd = sd(train$Seat.comfort),
  FlightDist_mean = mean(train$Flight.Distance), FlightDist_sd = sd(train$Flight.Distance),
  logDepDelay_mean = mean(logDepDelay), logDepDelay_sd = sd(logDepDelay),
  arr_delay_mean = mu_a, arr_delay_sd = sd_a
)

saveRDS(list(loo = loo_final, waic = waic_final,
             beta_mean = post_mean, beta_ci = post_ci, or_tbl = or_tbl,
             beta_draws = beta_draws, standardization = standardization),
        "section02_loo.rds")
```
