---
title: "2. Core Bayesian logistic regression (JAGS, vague priors)"
subtitle: "Airline Customer Satisfaction -- Bayesian Learning project"
output:
  html_document:
    df_print: paged
    mathjax: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. Data

We start from the same stratified training set used in `01_conjugate_baseline.Rmd`.

```{r libraries}
library(rjags)
library(coda)
```

```{r load-data}
train <- read.csv("data/train_subsample.csv", stringsAsFactors = FALSE)
train$cabin_class <- factor(train$cabin_class, levels = c("Eco", "Eco Plus", "Business"))

y <- as.numeric(train$satisfaction == "satisfied")
n <- length(y)

cat("n =", n, " | satisfied =", sum(y), "\n")
```

## 1. Model 0: intercept only

Before adding any covariates, let's verify JAGS environment is working
with the simplest model possible

$$y_i \mid p \overset{ind}{\sim} \text{Bernoulli}(p), \qquad \text{logit}(p) = \beta_0, \qquad \beta_0 \sim \mathcal{N}(0, 100)$$

```{r model-intercept-only}
model_string <- "
model {
  for (i in 1:n) {
  y[i] ~ dbern(p) 
  }
  logit(p) <- beta0
  beta0 ~ dnorm(0, 0.01)
}
"
```

## 2. Fit

```{r fit-intercept-only}
jags_data <- list(y = y, n = n)

set.seed(2026)
jm <- jags.model(textConnection(model_string), data = jags_data, n.chains = 4, n.adapt = 1000)

update(jm, n.iter = 1000)

samples <- coda.samples(jm, variable.names = c("beta0", "p"), n.iter = 5000)
```

## 3. Diagnostics

```{r diagnostics-intercept-only}
plot(samples[, "beta0"])
gelman.diag(samples)
effectiveSize(samples)

# Save the intercept trace + density for the report appendix. beta0 is usually
# the slowest-mixing parameter in a regression, so it is the one worth showing.
dir.create("figs", showWarnings = FALSE)
png("figs/trace_intercept.png", width = 8, height = 3.2, units = "in", res = 150)
plot(samples[, "beta0"])
invisible(dev.off())
```

## 4. Comparison with the conjugate baseline (Section 1)

As a check, we compare the intercept-only model's implied satisfaction rate, ilogit(β̂₀), against the closed-form Beta(1,1) posterior mean from Section 1. With $n \approx 1200$ the two priors barely differ, so the numbers should match up to Monte Carlo error. If they don't, the problem is in the JAGS code and not in the model.

```{r compare-with-section1}
x_baseline <- sum(y)
p_hat_baseline <- (1 + x_baseline) / (2 + n)
p_hat_jags <- mean(plogis(as.matrix(samples)[, "beta0"]))

cat("Baseline (Beta-Binomial):", round(p_hat_baseline, 4), "\n")
cat("JAGS (intercept only):   ", round(p_hat_jags, 4), "\n")
cat("Absolute difference:     ", signif(abs(p_hat_baseline - p_hat_jags), 3), "\n")
```

The two agree to `r signif(abs(p_hat_baseline - p_hat_jags), 2)`, so the JAGS setup recovers the right baseline before we add any covariates.

## 5. Adding the covariates

We extend the baseline with a Bayesian logistic regression on seven covariates identified in the Section 0 exploratory analysis: age, seat comfort rating, log-transformed flight distance, cabin class (two dummies, Economy as reference), log-transformed departure delay, the decorrelated arrival-delay residual, travel type, and customer type. Continuous covariates are centered and scaled prior to fitting. We place independent weakly-informative N(0,100) priors on the intercept and all slope coefficients, estimated via JAGS (4 chains, 1000 adaptation iterations, 2000 burn-in, 2000 retained draws per chain).

```{r}
train$flight_distance_z <- as.numeric(scale(log1p(train$flight_distance)))
train$dep_delay_z       <- as.numeric(scale(log1p(train$dep_delay)))
train$arr_delay_extra_z <- as.numeric(scale(train$arr_delay_extra))  
train$age_z             <- as.numeric(scale(train$age))
train$seat_comfort_z    <- as.numeric(scale(train$seat_comfort))

X <- model.matrix(~ age_z + seat_comfort_z + flight_distance_z + cabin_class +
                     dep_delay_z + arr_delay_extra_z + travel_type + customer_type,
                   data = train)[, -1]   # -1 minus the intercept, beta0 alone

p_cov <- ncol(X)
colnames(X)
```
```{r}
model_string_cov <- "
model {
  for (i in 1:n) {
  y[i] ~ dbern(p[i])
  logit(p[i]) <- beta0 + inprod(X[i,], beta[])
  log_lik[i] <- logdensity.bern(y[i], p[i])
  }
  beta0 ~ dnorm(0, 0.01)
  for (j in 1:p_cov){
  beta[j] ~ dnorm(0, 0.01)
  }
}
"

jags_data_cov <- list (y = y, X = X, n = n, p_cov = p_cov)

set.seed(2026)
jm_cov <- jags.model(textConnection(model_string_cov), data = jags_data_cov, n.chains = 4, n.adapt = 1000)

update(jm_cov, n.iter = 2000)

samples_cov <- coda.samples(jm_cov, variable.names = c("beta0", "beta", "log_lik"),
                             n.iter = 2000)
```
```{r}
beta_cols <- c("beta0", paste0("beta[", 1:p_cov, "]"))
gelman.diag(samples_cov[, beta_cols])
effectiveSize(samples_cov[, beta_cols])
```
```{r}
summary(samples_cov[, beta_cols])
```


## 6. cabin_class encoding: dummy vs. ordinal

```{r}
train$cabin_class_ord <- as.integer(train$cabin_class)
train$cabin_class_ord_z <- as.numeric(scale(train$cabin_class_ord))

X_ord <- model.matrix(~ age_z + seat_comfort_z + flight_distance_z + cabin_class_ord_z +
                          dep_delay_z + arr_delay_extra_z + travel_type + customer_type,
                        data = train)[, -1]

p_cov_ord <- ncol(X_ord)
colnames(X_ord)
```
```{r}
jags_data_ord <- list(y = y, X = X_ord, n = n, p_cov = p_cov_ord)

set.seed(2026)
jm_ord <- jags.model(textConnection(model_string_cov), data = jags_data_ord,
                      n.chains = 4, n.adapt = 1000)   # model_string_cov is generic: it reads whatever X is passed in

update(jm_ord, n.iter = 2000)

samples_ord <- coda.samples(jm_ord, variable.names = c("beta0", "beta", "log_lik"),
                             n.iter = 2000)
```
```{r}
library(loo)

extract_ll <- function(samples) {
  m <- as.matrix(samples)
  m[, grep("^log_lik\\[", colnames(m))]
}

waic_cov <- waic(extract_ll(samples_cov))   # Section 5: dummy encoding
waic_ord <- waic(extract_ll(samples_ord))   # Section 6: ordinal encoding

# Named list so the comparison table labels the rows, rather than model1/model2
loo_compare(list(dummy = waic_cov, ordinal = waic_ord))
```
Comparing the two cabin-class encodings via WAIC, the dummy-coded specification is preferred over the ordinal score (elpd_diff = 6.4, SE = 3.5), consistent with the coefficient pattern in Section 5: satisfaction jumps sharply for Business class but does not differ meaningfully between Economy and Economy Plus, violating the equal-spacing assumption an ordinal encoding imposes. The preference is moderate rather than decisive (elpd_diff/se_diff ≈ 1.8), so we retain the dummy encoding as the primary specification.

## 7. Link function: logit vs. probit

```{r}
model_string_probit <- "
model {
  for (i in 1:n) {
    y[i] ~ dbern(p[i])
    probit(p[i]) <- beta0 + inprod(X[i,], beta[])
    log_lik[i] <- logdensity.bern(y[i], p[i])
  }
  beta0 ~ dnorm(0, 0.01)
  for (j in 1:p_cov) {
    beta[j] ~ dnorm(0, 0.01)
  }
}
"

set.seed(2026)
jm_probit <- jags.model(textConnection(model_string_probit), data = jags_data_cov,
                         n.chains = 4, n.adapt = 1000)

update(jm_probit, n.iter = 2000)

samples_probit <- coda.samples(jm_probit, variable.names = c("beta0", "beta", "log_lik"),
                                n.iter = 2000)
```

```{r}
waic_probit <- waic(extract_ll(samples_probit))

loo_compare(list(logit = waic_cov, probit = waic_probit))
```
Comparing the logit and probit link functions via WAIC yields an elpd_diff of only -0.8 (SE = 0.6) — well within the range where loo's own diagnostics flag the difference as too small to distinguish reliably. The two link functions are therefore predictively equivalent for this specification, as expected given their near-identical shape; we retain the logit link going forward for its more direct interpretability (coefficients read as log-odds).
