---
title: "04 - Cross-check: BAS Models"
subtitle: "Airline Customer Satisfaction -- Bayesian Learning & Monte Carlo Simulations, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                       fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(BAS)
```

Two models, two purposes: (a) a deliberately naive linear-probability fit
of `satisfaction` via `bas.lm`, to show concretely why it's the wrong
likelihood for a binary outcome; (b) `bas.glm` with a binomial family,
the natural GLM extension of the `bas.lm` workflow from the course --
g-prior / BIC logic, Bernoulli model prior, posterior inclusion
probabilities, BMA. This section answers the brief's own question
("BAS or JAGS?") with "both, for different jobs": Section 03 gives one
model with full shrinkage; this section asks whether a completely
different estimation approach -- stochastic search over the *model
space* itself, rather than shrinkage within one model -- lands on the
same answer.

```{r load-data}
train <- read.csv("airline_train_n1200.csv", stringsAsFactors = FALSE)
train$y <- as.integer(train$satisfaction == "satisfied")

std <- function(x) as.numeric(scale(x))

service_items <- c("Seat.comfort", "Departure.Arrival.time.convenient", "Food.and.drink",
                    "Gate.location", "Inflight.wifi.service", "Inflight.entertainment",
                    "Online.support", "Ease.of.Online.booking", "On.board.service",
                    "Leg.room.service", "Baggage.handling", "Checkin.service",
                    "Cleanliness", "Online.boarding")
X_service <- apply(train[, service_items], 2, std)

logDepDelay  <- log1p(train$Departure.Delay.in.Minutes)
signedLogArr <- sign(train$ArrivalDelay.Residual) * log1p(abs(train$ArrivalDelay.Residual))

Age_std         <- std(train$Age)
FlightDist_std  <- std(train$Flight.Distance)
logDepDelay_std <- std(logDepDelay)

obs  <- !is.na(signedLogArr)
mu_a <- mean(signedLogArr[obs]); sd_a <- sd(signedLogArr[obs])
arr_delay_std <- (signedLogArr - mu_a) / sd_a

Class_EcoPlus  <- as.integer(train$Class == "Eco Plus")
Class_Business <- as.integer(train$Class == "Business")
TypeTravel_Personal <- as.integer(train$Type.of.Travel == "Personal Travel")
CustType_disloyal   <- as.integer(train$Customer.Type == "disloyal Customer")

X_extra <- cbind(Age_std, FlightDist_std, Class_EcoPlus, Class_Business,
                  logDepDelay_std, TypeTravel_Personal, CustType_disloyal)
X_full <- cbind(X_service, X_extra)
```

`bas.lm`/`bas.glm` cannot use JAGS's stochastic-imputation trick for the
one missing arrival-delay residual, so for this section only we drop
that single row (negligible: 1 of 1203).

```{r drop-missing}
keep <- !is.na(arr_delay_std)
X_bas <- cbind(X_full[keep, ], ArrivalDelay_resid = arr_delay_std[keep])
y_bas <- train$y[keep]
df_bas <- data.frame(y = y_bas, X_bas)

cat("Dropped", sum(!keep), "row(s); df_bas dim:", dim(df_bas), "\n")
```

# (a) Linear probability model -- deliberately the wrong likelihood

```{r fit-lin, cache=TRUE}
set.seed(123)
fit_lin <- bas.lm(y ~ ., data = df_bas, prior = "BIC", modelprior = uniform(),
                   method = "MCMC", MCMC.iterations = 50000)
pred_lin <- predict(fit_lin, estimator = "BMA")$fit
```

```{r lin-plot, fig.width=6, fig.height=3.5}
data.frame(pred = pred_lin) %>%
  ggplot(aes(pred)) +
  geom_histogram(bins = 40, fill = "#D85A30", alpha = 0.8) +
  geom_vline(xintercept = c(0, 1), linetype = "dashed", color = "black") +
  labs(title = "bas.lm treats satisfaction as Gaussian",
       subtitle = sprintf("%.1f%% of fitted probabilities fall outside [0,1]",
                           100 * mean(pred_lin < 0 | pred_lin > 1)),
       x = "Predicted P(satisfied)", y = "Count") +
  theme_minimal(base_size = 11)
```

`r sprintf("%.1f%%", 100*mean(pred_lin < 0 | pred_lin > 1))` of the
in-sample fitted values fall outside `[0,1]` -- not a rounding curiosity,
but the direct, visible consequence of using a likelihood that assumes
constant variance and an unbounded response for what is actually a
bounded 0/1 outcome. This is the concrete justification for Sections 02
and 03, and for part (b) below.

# (b) `bas.glm` with a binomial family

With `r ncol(X_full)+1` predictors, $2^{`r ncol(X_full)+1`} \approx$
`r format(round(2^(ncol(X_full)+1)/1e6,1), big.mark=",")` million
candidate models -- too many to enumerate exhaustively, so we use
`method = "MCMC"` to stochastically search the model space, as flagged
in Section 00's instability discussion.

```{r fit-glm-g, cache=TRUE}
set.seed(123)
fit_glm_g <- bas.glm(y ~ ., data = df_bas, family = binomial(),
                      betaprior = g.prior(nrow(df_bas)), modelprior = uniform(),
                      method = "MCMC", MCMC.iterations = 100000)
```

```{r top-models}
knitr::kable(round(summary(fit_glm_g, n.models = 5), 3),
             caption = "Top 5 models by posterior probability (g-prior)")
```

The single best model
(posterior probability `r sprintf("%.3f", max(fit_glm_g$postprobs))`)
includes `Departure/Arrival time convenient`, `Inflight entertainment`,
`Ease of Online booking`, `On-board service`, `Leg room service`,
`Checkin service`, `Personal Travel` and `disloyal Customer` -- already
visibly the same handful of variables the horseshoe in Section 03
preserved.

# Is the MCMC search actually stable?

This is exactly the instability question flagged back in Section 00:
with ~22 correlated candidate predictors, a stochastic search over
$4.2$ million models could plausibly land on different answers between
runs. We check this three ways rather than assuming it away.

**Convergence within one run.** BAS reports two independent estimates of
each posterior inclusion probability: one from renormalizing the
sampled models' marginal likelihoods, one from raw MCMC visit
frequency. If the chain has converged, they should agree closely.

```{r diagnostics-plot, fig.width=5, fig.height=5}
diag_df <- data.frame(renormalized = fit_glm_g$probne0.RN,
                       mcmc = fit_glm_g$probne0.MCMC)
ggplot(diag_df, aes(renormalized, mcmc)) +
  geom_point(color = "#185FA5", alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  coord_equal(xlim = c(0,1), ylim = c(0,1)) +
  labs(title = "MCMC convergence check: PIP estimates should lie on the diagonal",
       x = "PIP (renormalized marginal likelihood)", y = "PIP (raw MCMC frequency)") +
  theme_minimal(base_size = 11)
```

Correlation between the two estimates is
`r sprintf("%.4f", cor(fit_glm_g$probne0.RN, fit_glm_g$probne0.MCMC))`,
maximum absolute difference
`r sprintf("%.3f", max(abs(fit_glm_g$probne0.RN - fit_glm_g$probne0.MCMC)))`
-- the points sit essentially on the diagonal, the sign this run has
converged.

**Stability across seeds.** We refit the identical model twice more with
different random seeds and compare posterior inclusion probabilities.

```{r seed-stability, cache=TRUE}
run_seed <- function(seed) {
  set.seed(seed)
  fit <- bas.glm(y ~ ., data = df_bas, family = binomial(),
                 betaprior = g.prior(nrow(df_bas)), modelprior = uniform(),
                 method = "MCMC", MCMC.iterations = 100000)
  fit$probne0
}
pip_seed1 <- run_seed(1)
pip_seed2 <- run_seed(2)

pip_df <- data.frame(term = fit_glm_g$namesx,
                      seed123 = fit_glm_g$probne0,
                      seed1 = pip_seed1, seed2 = pip_seed2)
pip_df$range <- apply(pip_df[, 2:4], 1, function(x) max(x) - min(x))
```

```{r seed-stability-table}
knitr::kable(pip_df %>% arrange(desc(range)) %>% head(8), row.names = FALSE, digits = 3,
             caption = "Largest PIP swings across 3 seeds (of 22 predictors)")
cat(sprintf("Max PIP range across seeds: %.3f\nMean PIP range across seeds: %.4f\n",
            max(pip_df$range), mean(pip_df$range)))
```

**BIC prior as a robustness check.** A different prior on the regression
coefficients is a different, non-nested check on the same question.

```{r bic-prior, cache=TRUE}
set.seed(123)
fit_glm_bic <- bas.glm(y ~ ., data = df_bas, family = binomial(),
                        betaprior = bic.prior(nrow(df_bas)), modelprior = uniform(),
                        method = "MCMC", MCMC.iterations = 100000)

cmp_prior <- data.frame(term = fit_glm_g$namesx, g_prior = round(fit_glm_g$probne0, 3),
                         BIC = round(fit_glm_bic$probne0, 3))
cmp_prior$abs_diff <- round(abs(cmp_prior$g_prior - cmp_prior$BIC), 3)
```

```{r bic-table}
knitr::kable(cmp_prior %>% arrange(desc(abs_diff)) %>% head(6), row.names = FALSE,
             caption = "Largest g-prior vs. BIC-prior PIP differences")
cat("Max |g-prior - BIC| difference:", max(cmp_prior$abs_diff), "\n")
```

With 100,000 MCMC iterations, PIPs move by at most
`r sprintf("%.3f", max(pip_df$range))` across three different seeds and
by at most `r sprintf("%.3f", max(cmp_prior$abs_diff))` between the
g-prior and the BIC-based prior. **This directly answers Section 00's
instability question**: the "instability" earlier runs might show is a
symptom of too few MCMC iterations relative to a large, correlated model
space -- not an inherent flaw of the g-prior or of BAS. With an adequate
iteration budget the search is, in practice, stable.

# BAS vs. the horseshoe: do two independent methods agree?

```{r hs-comparison}
hs <- readRDS("horseshoe_summary.rds")
hs_names_raw <- c("Seat.comfort", "Departure.Arrival.time.convenient", "Food.and.drink",
                   "Gate.location", "Inflight.wifi.service", "Inflight.entertainment",
                   "Online.support", "Ease.of.Online.booking", "On.board.service",
                   "Leg.room.service", "Baggage.handling", "Checkin.service",
                   "Cleanliness", "Online.boarding", "Age_std", "FlightDist_std",
                   "Class_EcoPlus", "Class_Business", "logDepDelay_std",
                   "TypeTravel_Personal", "CustType_disloyal", "ArrivalDelay_resid")
kappa_ordered <- hs$post_kappa[c(paste0("beta[", 1:21, "]"), "beta_arr")]

cmp_methods <- data.frame(term = hs_names_raw,
                           PIP = round(fit_glm_g$probne0[-1], 3),
                           kappa = round(kappa_ordered, 3))
cmp_methods$signal_BAS <- cmp_methods$PIP > 0.5
cmp_methods$signal_HS  <- cmp_methods$kappa < 0.5
cmp_methods$agree <- cmp_methods$signal_BAS == cmp_methods$signal_HS
agreement_rate <- mean(cmp_methods$agree)
```

```{r hs-comparison-plot, fig.width=7, fig.height=6}
cmp_methods %>%
  mutate(term = factor(term, levels = term[order(PIP)])) %>%
  ggplot(aes(x = PIP, y = 1 - kappa, color = agree)) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey60") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60") +
  scale_color_manual(values = c("TRUE" = "#3B6D11", "FALSE" = "#D85A30"),
                      labels = c("TRUE" = "Agree", "FALSE" = "Disagree"), name = NULL) +
  labs(title = "BAS posterior inclusion probability vs. horseshoe (1 - kappa)",
       subtitle = sprintf("%.0f%% agreement on signal (>0.5) vs. noise (<=0.5)", 100*agreement_rate),
       x = "PIP (BAS, g-prior)", y = expression(1 - kappa[j]~"(horseshoe)")) +
  theme_minimal(base_size = 10)
```

`r sprintf("%.0f%%", 100*agreement_rate)` of the 22 predictors get the
same signal/noise call from BAS's model-space search and from the
horseshoe's continuous shrinkage -- two structurally different
estimation strategies (discrete model averaging vs. a single shrinkage
prior) landing on the same handful of variables. The one disagreement,
`Class: Business` (PIP `r cmp_methods$PIP[cmp_methods$term=="Class_Business"]`,
$\kappa=$ `r cmp_methods$kappa[cmp_methods$term=="Class_Business"]`), is a
genuine borderline case in *both* methods (BAS just below its 0.5 cutoff,
the horseshoe's credible interval for it crossing zero in Section 03) --
not a real contradiction between them.

# Export for later sections

Saved for reuse in Section 06: the fitted `bas.glm` object itself (BAS's
`predict()` method supports genuine BMA prediction on new data, so no
separate coefficient-draw export is needed the way JAGS required). The
covariate construction here is identical to Section 03's, so the same
training-set standardization parameters already saved in
`section03_loo.rds` apply -- no need to duplicate them.

```{r export-bas}
saveRDS(fit_glm_g, "section04_basfit.rds")
```
