---
title: "Section 5 — WAIC/PSIS-LOO Comparison and Prior Sensitivity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Motivation

Section 2 fit a flat-prior logistic model on 9 predictors; Section 3 fit a
horseshoe-shrinkage logistic model on the full 22-predictor set. The
coefficient plots in those sections show *what* each model estimates, but
not *which model actually predicts the training data better once complexity
is penalized*, or *how much the flat model's conclusions depend on the
specific prior variance chosen* ($\sigma^2 = 100$) rather than being an
artifact of that choice. This section answers both questions via WAIC,
PSIS-LOO, and a prior-variance sensitivity sweep.

```{r setup-packages}
rm(list = ls())
set.seed(2026)

library(rjags)
library(coda)
library(loo)
library(dplyr)
library(ggplot2)

load.module("dic")   # enables logdensity.bern() inside the JAGS model string
```

```{r load-data}
train <- read.csv("data/train_subsample.csv", stringsAsFactors = FALSE)
names(train) <- make.names(names(train), unique = TRUE)

train$satisfaction  <- factor(train$satisfaction, levels = c("dissatisfied", "satisfied"))
train$customer_type <- factor(train$customer_type)
train$travel_type   <- factor(train$travel_type)
train$cabin_class   <- factor(train$cabin_class, levels = c("Eco", "Eco Plus", "Business"))

train$flight_distance_z <- as.numeric(scale(log1p(train$flight_distance)))
train$dep_delay_z       <- as.numeric(scale(log1p(train$dep_delay)))
train$arr_delay_extra_z <- as.numeric(scale(train$arr_delay_extra))
train$age_z             <- as.numeric(scale(train$age))
train$seat_comfort_z    <- as.numeric(scale(train$seat_comfort))

y <- as.numeric(train$satisfaction == "satisfied")
n <- length(y)

X_flat <- model.matrix(~ age_z + seat_comfort_z + flight_distance_z + cabin_class +
                          dep_delay_z + arr_delay_extra_z + travel_type + customer_type,
                        data = train)[, -1]
p_flat <- ncol(X_flat)
```

## Flat model, refit with log-likelihood monitored

`loo::waic()`/`loo::loo()` need the full $n \times S$ matrix of
per-observation, per-draw log-likelihood, not a single deviance summary. The
Section 2 refit only tracked `beta0, beta`, so the flat model is refit here
with `log_lik` added via JAGS's `logdensity.bern()`.

```{r flat-refit-loglik}
model_string_flat <- "
model {
  for (i in 1:n) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- beta0 + inprod(X[i,], beta[])
    log_lik[i] <- logdensity.bern(y[i], p[i])
  }
  beta0 ~ dnorm(0, 0.01)
  for (j in 1:p_flat) { beta[j] ~ dnorm(0, 0.01) }
}
"

jm_flat <- jags.model(textConnection(model_string_flat),
                       data = list(y = y, X = X_flat, n = n, p_flat = p_flat),
                       n.chains = 4, n.adapt = 1000)
update(jm_flat, n.iter = 2000)
samples_flat <- coda.samples(jm_flat, variable.names = c("beta0", "beta", "log_lik"),
                              n.iter = 2000)
flat_mat <- as.matrix(samples_flat)

loglik_flat <- flat_mat[, grep("^log_lik\\[", colnames(flat_mat))]
```

## Horseshoe: reuse saved log-likelihood

`log_lik` was already monitored in Section 3's `variable.names` and saved to
`data/samples_hs.rds`, so it only needs to be loaded and extracted here, not
recomputed.

```{r load-horseshoe}
samples_hs <- readRDS("data/samples_hs.rds")
samp_mat   <- as.matrix(samples_hs)

loglik_hs <- samp_mat[, grep("^log_lik\\[", colnames(samp_mat))]
```

## WAIC and PSIS-LOO comparison

```{r waic-loo}
waic_flat <- loo::waic(loglik_flat)
waic_hs   <- loo::waic(loglik_hs)

loo_flat <- loo::loo(loglik_flat)
loo_hs   <- loo::loo(loglik_hs)

print(waic_flat)
print(waic_hs)

comparison <- loo::loo_compare(list(flat = loo_flat, horseshoe = loo_hs))
print(comparison)
```

The horseshoe model fits the training data decisively better once complexity
is penalized: WAIC is `r round(waic_flat$estimates["waic","Estimate"],1)`
(SE `r round(waic_flat$estimates["waic","SE"],1)`) for the flat model versus
`r round(waic_hs$estimates["waic","Estimate"],1)`
(SE `r round(waic_hs$estimates["waic","SE"],1)`) for the horseshoe, and
`loo_compare` puts the flat model
`r round(abs(comparison["flat","elpd_diff"]),1)` ELPD points behind the
horseshoe (SE `r round(comparison["flat","se_diff"],1)`) — an order of
magnitude beyond noise.

This is **not**, however, a clean test of "horseshoe prior vs. flat prior"
holding everything else equal. The flat model does not just have a different
prior; it has 13 fewer predictors. It never sees entertainment, wifi, food,
or the other service-quality items, several of which Section 3's $\kappa_j$
plot flagged as real signal. A substantial share of this WAIC gap is
plausibly attributable to "a model with more of the right variables fits
better," not to shrinkage priors outperforming flat priors per se — worth
stating explicitly rather than implying the prior alone did this work.

A second honest note: `p_waic` for the horseshoe
(`r round(waic_hs$estimates["p_waic","Estimate"],1)`) sits close to its 23
nominal parameters, much as the flat model's
(`r round(waic_flat$estimates["p_waic","Estimate"],1)`) sits close to its 10
nominal parameters — despite the real per-coefficient shrinkage visible in
the $\kappa_j$ plot. `p_waic` is a variance-based approximation of
*effective* complexity; it need not drop just because individual
coefficients are pulled toward zero, especially when posterior uncertainty
in the tails remains real.

## Pareto k diagnostics

```{r pareto-k}
k_flat <- loo_flat$diagnostics$pareto_k
k_hs   <- loo_hs$diagnostics$pareto_k

cat(sprintf("Flat model: %d / %d observations with Pareto k > 0.7\n",
            sum(k_flat > 0.7), length(k_flat)))
cat(sprintf("Horseshoe model: %d / %d observations with Pareto k > 0.7\n",
            sum(k_hs > 0.7), length(k_hs)))

plot(loo_hs, main = "PSIS-LOO Pareto k diagnostic — horseshoe model")
```

`r sum(k_hs > 0.7)` of `r length(k_hs)` horseshoe observations and
`r sum(k_flat > 0.7)` of `r length(k_flat)` flat-model observations exceed
the $k > 0.7$ threshold where PSIS-LOO becomes unreliable. All Pareto $k$
values for the horseshoe model sit comfortably under 0.4 (see the diagnostic
plot), well inside the "good" range. PSIS-LOO is fully trustworthy for both
models here, so the ELPD comparison above can be read at face value, with no
reliability caveat needed.

## Prior sensitivity analysis (flat model, $\sigma^2 \in \{1, 10, 100, 1000\}$)

The horseshoe already estimates its own sparsity from the data, so there is
no single fixed prior variance to sweep there. The check below instead
targets the flat model's fixed $\mathcal N(0,\sigma^2)$ prior on $\beta$,
refitting at four variances and tracking both the coefficient estimates and
WAIC. Note $\sigma^2$ is a *variance*, but JAGS's `dnorm()` takes
*precision* as its second argument — the loop converts via `prec <- 1 / s2`
at each step.

```{r sensitivity-sweep}
sigma2_grid <- c(1, 10, 100, 1000)

sens_beta_list <- list()
sens_waic_list <- list()

for (s2 in sigma2_grid) {
  prec <- 1 / s2

  model_string_sens <- sprintf("
  model {
    for (i in 1:n) {
      y[i] ~ dbern(p[i])
      logit(p[i]) <- beta0 + inprod(X[i,], beta[])
      log_lik[i] <- logdensity.bern(y[i], p[i])
    }
    beta0 ~ dnorm(0, %f)
    for (j in 1:p_flat) { beta[j] ~ dnorm(0, %f) }
  }", prec, prec)

  jm_s <- jags.model(textConnection(model_string_sens),
                      data = list(y = y, X = X_flat, n = n, p_flat = p_flat),
                      n.chains = 4, n.adapt = 1000, quiet = TRUE)
  update(jm_s, n.iter = 2000, progress.bar = "none")
  samp_s <- coda.samples(jm_s, variable.names = c("beta0", "beta", "log_lik"),
                          n.iter = 2000, progress.bar = "none")
  mat_s <- as.matrix(samp_s)

  beta_cols <- c("beta0", paste0("beta[", seq_len(p_flat), "]"))
  sens_beta_list[[as.character(s2)]] <- data.frame(
    sigma2    = s2,
    Predictor = c("(Intercept)", colnames(X_flat)),
    Mean      = colMeans(mat_s[, beta_cols])
  )

  loglik_s <- mat_s[, grep("^log_lik\\[", colnames(mat_s))]
  sens_waic_list[[as.character(s2)]] <- data.frame(
    sigma2 = s2,
    waic   = loo::waic(loglik_s)$estimates["waic", "Estimate"]
  )
}

sens_beta_df <- bind_rows(sens_beta_list)
sens_waic_df <- bind_rows(sens_waic_list)

print(sens_waic_df)
```

```{r sensitivity-plots}
waic_se  <- waic_flat$estimates["waic", "SE"]
waic_est <- waic_flat$estimates["waic", "Estimate"]

p_sens_waic <- ggplot(sens_waic_df, aes(x = factor(sigma2), y = waic, group = 1)) +
  geom_hline(yintercept = waic_est, linetype = "dotted") +
  geom_ribbon(aes(ymin = waic_est - waic_se, ymax = waic_est + waic_se),
              alpha = 0.15, fill = "steelblue") +
  geom_line() + geom_point(size = 2) +
  labs(title = "WAIC across prior variances (flat model)",
       subtitle = "Shaded band = \u00b11 SE of the WAIC estimate",
       x = expression(sigma^2), y = "WAIC") +
  theme_minimal()
ggsave("figs/prior_sensitivity_waic.png", p_sens_waic, width = 6, height = 4, dpi = 150)
p_sens_waic

p_sens_beta <- ggplot(sens_beta_df, aes(x = factor(sigma2), y = Mean,
                                          group = Predictor, color = Predictor)) +
  geom_line() + geom_point() +
  labs(title = "Posterior mean coefficients across prior variances (flat model)",
       x = expression(sigma^2 ~ "(prior variance on" ~ beta ~ ")"),
       y = expression(beta[j])) +
  theme_minimal()
ggsave("figs/prior_sensitivity_beta.png", p_sens_beta, width = 8, height = 5, dpi = 150)
p_sens_beta
```

```{r beta-stability}
beta_range <- sens_beta_df %>%
  group_by(Predictor) %>%
  summarise(range = diff(range(Mean))) %>%
  arrange(desc(range))
print(beta_range)
```

WAIC ranges from `r round(min(sens_waic_df$waic),2)` to
`r round(max(sens_waic_df$waic),2)` across $\sigma^2 \in \{1,10,100,1000\}$
— a spread of `r round(diff(range(sens_waic_df$waic)),2)`, negligible next
to the flat model's own WAIC standard error of `r round(waic_se,1)`. The
largest posterior-mean coefficient shift across the same range is
`r round(beta_range$range[1],3)`, for `r beta_range$Predictor[1]`. Both the
WAIC plot's shaded $\pm 1$ SE band and the coefficient-stability plot
confirm the flat model's conclusions are effectively prior-insensitive
across the tested range, which supports $\sigma^2 = 100$ (used throughout
Sections 2–3) as a safe default rather than an arbitrary choice.

## Summary

WAIC and PSIS-LOO favor the horseshoe decisively, though part of that gap
reflects its larger predictor set rather than the shrinkage prior alone —
worth stating plainly rather than overclaiming. Pareto $k$ diagnostics
confirm PSIS-LOO is reliable for both models, so the comparison can be
trusted at face value. The prior sensitivity sweep shows the flat model's
coefficient estimates and WAIC are stable across four orders of magnitude in
$\sigma^2$, justifying $\sigma^2 = 100$ as a safe rather than arbitrary
choice throughout the project.
