---
title: "Section 6 — Prediction and Validation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Motivation

Sections 2–4 fit three models on the training subsample — a flat-prior JAGS
logistic regression, a horseshoe-shrinkage JAGS logistic regression, and
BAS's `bas.glm` — and Section 5 showed the horseshoe fits the *training*
data decisively better once complexity is penalized (WAIC/PSIS-LOO). None
of the three has touched the held-out test set yet, so none of that is a
genuine test of generalization. This section is that test.

For each test customer, the predicted probability is the **posterior mean**
of $\text{logit}^{-1}(\beta_0 + x_i^\top\beta)$, averaged over posterior
draws — not $\text{logit}^{-1}$ evaluated once at the posterior mean of
$\beta$. These differ because the logit link is nonlinear
($\mathbb E[g(\theta)] \neq g(\mathbb E[\theta])$ for nonlinear $g$), and
only the draw-averaged version is the actual posterior predictive mean.
Predictions are classified at a 0.5 cutoff and reported as accuracy and a
confusion matrix.



```{r setup-packages}
rm(list = ls())
set.seed(2026)

library(rjags)
library(coda)
library(BAS)
```

## Load train and test, align factor levels

Test must reuse train's exact factor levels. A different level ordering
changes `model.matrix()`'s dummy coding — same variable, different columns
— and every coefficient from Sections 2–4 would get silently misapplied.

```{r load-data}
train <- read.csv("data/train_subsample.csv", stringsAsFactors = FALSE)
names(train) <- make.names(names(train), unique = TRUE)

test <- read.csv("data/test_holdout.csv", stringsAsFactors = FALSE)
names(test) <- make.names(names(test), unique = TRUE)

stopifnot(all(names(train) %in% names(test)))

train$satisfaction  <- factor(train$satisfaction, levels = c("dissatisfied", "satisfied"))
train$customer_type <- factor(train$customer_type)
train$travel_type   <- factor(train$travel_type)
train$cabin_class   <- factor(train$cabin_class, levels = c("Eco", "Eco Plus", "Business"))

test$satisfaction  <- factor(test$satisfaction,  levels = levels(train$satisfaction))
test$customer_type <- factor(test$customer_type, levels = levels(train$customer_type))
test$travel_type   <- factor(test$travel_type,   levels = levels(train$travel_type))
test$cabin_class   <- factor(test$cabin_class,   levels = levels(train$cabin_class))

n_train <- nrow(train)
n_test  <- nrow(test)
```

## Standardize test using train's center/scale

The held-out set's own mean/sd never touches model fitting or evaluation —
reusing train's center/scale is what makes this an honest leakage-free test.

```{r standardize}
standardize_with_train <- function(train_raw, test_raw) {
  ctr <- mean(train_raw); scl <- sd(train_raw)
  list(train_z = (train_raw - ctr) / scl, test_z = (test_raw - ctr) / scl)
}

fd  <- standardize_with_train(log1p(train$flight_distance), log1p(test$flight_distance))
dd  <- standardize_with_train(log1p(train$dep_delay),       log1p(test$dep_delay))
ade <- standardize_with_train(train$arr_delay_extra,        test$arr_delay_extra)
ag  <- standardize_with_train(train$age,                    test$age)
sc  <- standardize_with_train(train$seat_comfort,            test$seat_comfort)

train$flight_distance_z <- fd$train_z; test$flight_distance_z <- fd$test_z
train$dep_delay_z       <- dd$train_z; test$dep_delay_z       <- dd$test_z
train$arr_delay_extra_z <- ade$train_z; test$arr_delay_extra_z <- ade$test_z
train$age_z             <- ag$train_z; test$age_z             <- ag$test_z
train$seat_comfort_z    <- sc$train_z; test$seat_comfort_z    <- sc$test_z

extra_service_items <- c("time_convenient", "food_drink", "gate_location", "wifi",
                          "entertainment", "online_support", "ease_booking",
                          "onboard_service", "legroom", "baggage", "checkin",
                          "cleanliness", "online_boarding")

for (v in extra_service_items) {
  sv <- standardize_with_train(train[[v]], test[[v]])
  train[[paste0(v, "_z")]] <- sv$train_z
  test[[paste0(v, "_z")]]  <- sv$test_z
}
extra_service_z <- paste0(extra_service_items, "_z")
```

## Design matrices

Same formulas as Sections 2 and 3, applied to both train and test, with a
hard check that the resulting column names line up exactly before either
matrix touches a model.

```{r design-matrices}
flat_formula <- ~ age_z + seat_comfort_z + flight_distance_z + cabin_class +
                    dep_delay_z + arr_delay_extra_z + travel_type + customer_type

X_train_flat <- model.matrix(flat_formula, data = train)[, -1]
X_test_flat  <- model.matrix(flat_formula, data = test)[, -1]
stopifnot(identical(colnames(X_train_flat), colnames(X_test_flat)))

hs_formula <- as.formula(paste(
  "~ age_z + seat_comfort_z + flight_distance_z + cabin_class +",
  "dep_delay_z + arr_delay_extra_z + travel_type + customer_type +",
  paste(extra_service_z, collapse = " + ")
))

X_train_hs <- model.matrix(hs_formula, data = train)[, -1]
X_test_hs  <- model.matrix(hs_formula, data = test)[, -1]
stopifnot(identical(colnames(X_train_hs), colnames(X_test_hs)))

y_train <- as.numeric(train$satisfaction == "satisfied")
y_test  <- as.numeric(test$satisfaction  == "satisfied")
```

## Posterior draws: flat (fresh refit), horseshoe and BAS (reloaded)

The flat model is refit here without `log_lik` (not needed for prediction,
and it's the expensive part of the Section 5 version — dropping it keeps
this fast). The horseshoe posterior and the BAS fit are reused directly from
the `.rds` files Sections 3 and 4 already saved.

```{r fit-flat, cache=TRUE}
p_flat <- ncol(X_train_flat)

model_string_flat <- "
model {
  for (i in 1:ntr) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- beta0 + inprod(X[i,], beta[])
  }
  beta0 ~ dnorm(0, 0.01)
  for (j in 1:p_flat) { beta[j] ~ dnorm(0, 0.01) }
}
"
jm_flat <- jags.model(textConnection(model_string_flat),
                       data = list(y = y_train, X = X_train_flat, ntr = n_train, p_flat = p_flat),
                       n.chains = 4, n.adapt = 1000)
update(jm_flat, n.iter = 2000)
samples_flat <- coda.samples(jm_flat, variable.names = c("beta0", "beta"), n.iter = 2000)
flat_mat <- as.matrix(samples_flat)

beta0_flat <- flat_mat[, "beta0"]
beta_flat  <- flat_mat[, paste0("beta[", seq_len(p_flat), "]")]
```

```{r load-hs-bas}
samples_hs <- readRDS("data/samples_hs.rds")
hs_mat     <- as.matrix(samples_hs)
p_hs       <- ncol(X_train_hs)

beta0_hs <- hs_mat[, "beta0"]
beta_hs  <- hs_mat[, paste0("beta[", seq_len(p_hs), "]")]

bas_glm_fit <- readRDS("data/bas_glm_fit.rds")
```

## Posterior-mean predicted probability, averaged over draws

Done in batches over posterior draws rather than all at once — at
`r format(n_test, big.mark = ",")` test rows, materializing a full
`n_test`$\times S$ matrix for the flat model's 8,000 draws would run into
several GB in one allocation. Batching keeps memory bounded regardless of
how many draws either chain has.

```{r predict-function}
posterior_mean_prob <- function(X_test, beta0_samples, beta_samples, batch_size = 200) {
  n_test <- nrow(X_test)
  S <- length(beta0_samples)
  prob_sum <- numeric(n_test)
  starts <- seq(1, S, by = batch_size)
  for (s0 in starts) {
    s1  <- min(s0 + batch_size - 1, S)
    eta <- outer(rep(1, n_test), beta0_samples[s0:s1]) +
             X_test %*% t(beta_samples[s0:s1, , drop = FALSE])
    prob_sum <- prob_sum + rowSums(plogis(eta))
  }
  prob_sum / S
}

p_flat_test <- posterior_mean_prob(X_test_flat, beta0_flat, beta_flat)
p_hs_test   <- posterior_mean_prob(X_test_hs,   beta0_hs,   beta_hs)

# predict.bas already does BMA-averaging internally — no manual draw loop needed here
bas_pred   <- predict(bas_glm_fit, newdata = test, type = "response", estimator = "BMA")
p_bas_test <- bas_pred$fit
```

## Evaluation: accuracy and confusion matrix at a 0.5 cutoff

```{r evaluate}
evaluate_predictions <- function(p_hat, y_true) {
  pred <- ifelse(p_hat >= 0.5, 1, 0)
  cm  <- table(Predicted = pred, Actual = y_true)
  acc <- mean(pred == y_true)
  list(confusion = cm, accuracy = acc)
}

res_flat <- evaluate_predictions(p_flat_test, y_test)
res_hs   <- evaluate_predictions(p_hs_test,   y_test)
res_bas  <- evaluate_predictions(p_bas_test,  y_test)

cat("Flat model:\n");      print(res_flat$confusion)
cat("\nHorseshoe model:\n"); print(res_hs$confusion)
cat("\nBAS bas.glm:\n");     print(res_bas$confusion)

summary_df <- data.frame(
  Model    = c("Flat (9 predictors)", "Horseshoe (22 predictors)", "BAS bas.glm"),
  Accuracy = c(res_flat$accuracy, res_hs$accuracy, res_bas$accuracy)
)
print(summary_df)
```

```{r interpretation-setup, include=FALSE}
acc_tab    <- summary_df[order(-summary_df$Accuracy), ]
best_model <- acc_tab$Model[1]
best_acc   <- acc_tab$Accuracy[1]
worst_acc  <- acc_tab$Accuracy[nrow(acc_tab)]
acc_spread <- 100 * (best_acc - worst_acc)
```

## Results

On the held-out test set of `r format(n_test, big.mark = ",")` customers —
untouched during fitting, standardization, or variable selection — accuracy
at a 0.5 cutoff is `r sprintf("%.1f%%", 100 * res_flat$accuracy)` for the
flat model, `r sprintf("%.1f%%", 100 * res_hs$accuracy)` for the horseshoe
model, and `r sprintf("%.1f%%", 100 * res_bas$accuracy)` for BAS's
`bas.glm`. The best-performing model on this held-out set is
**`r best_model`**, at `r sprintf("%.1f%%", 100 * best_acc)`; the full
spread across all three is `r sprintf("%.1f", acc_spread)` percentage
points.

This is worth reading against Section 5 rather than in isolation: WAIC and
PSIS-LOO there favored the horseshoe decisively *in-sample*, on the
training data it was fit to. Whether that advantage survives contact with
genuinely new customers is exactly what this section tests — and the
honest reading is whichever way the numbers land here, not an assumption
that the WAIC winner must also win the accuracy comparison. A model with
13 extra predictors has more room to overfit the training subsample even
under a shrinkage prior; if its test accuracy comes in close to or below
the flat model's despite the large WAIC gap, that gap says more about
in-sample fit than out-of-sample generalization, and the report should say
so plainly rather than picking the in-sample winner by default.

The confusion matrices above give the raw counts behind each accuracy
figure — worth a glance for whether errors lean toward false positives or
false negatives for any one model, even without computing precision/recall
formally, since the outcome here is close to balanced and a lopsided error
pattern would be the more interesting thing to flag qualitatively.
