---
title: "Section 6 — Prediction and Validation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Motivation

Sections 2–4 fit three models on the training subsample — a flat-prior JAGS
logistic regression, a horseshoe-shrinkage JAGS logistic regression, and
BAS's `bas.glm` — and Section 5 showed the horseshoe fits the *training*
data decisively better once complexity is penalized (WAIC/PSIS-LOO). None
of the three has touched the held-out test set yet, so none of that is a
genuine test of generalization. This section is that test.

For each test customer, the predicted probability is the **posterior mean**
of $\text{logit}^{-1}(\beta_0 + x_i^\top\beta)$, averaged over posterior
draws — not $\text{logit}^{-1}$ evaluated once at the posterior mean of
$\beta$. These differ because the logit link is nonlinear
($\mathbb E[g(\theta)] \neq g(\mathbb E[\theta])$ for nonlinear $g$), and
only the draw-averaged version is the actual posterior predictive mean.
Predictions are classified at a 0.5 cutoff and reported as accuracy and a
confusion matrix.



```{r setup-packages}
rm(list = ls())
set.seed(2026)

library(rjags)
library(coda)
library(BAS)
library(ggplot2)   # calibration plot
library(pROC)      # ROC / AUC
```

## Load train and test, align factor levels

Test must reuse train's exact factor levels. A different level ordering
changes `model.matrix()`'s dummy coding — same variable, different columns
— and every coefficient from Sections 2–4 would get silently misapplied.

```{r load-data}
train <- read.csv("data/train_subsample.csv", stringsAsFactors = FALSE)
names(train) <- make.names(names(train), unique = TRUE)

test <- read.csv("data/test_holdout.csv", stringsAsFactors = FALSE)
names(test) <- make.names(names(test), unique = TRUE)

stopifnot(all(names(train) %in% names(test)))

train$satisfaction  <- factor(train$satisfaction, levels = c("dissatisfied", "satisfied"))
train$customer_type <- factor(train$customer_type)
train$travel_type   <- factor(train$travel_type)
train$cabin_class   <- factor(train$cabin_class, levels = c("Eco", "Eco Plus", "Business"))

test$satisfaction  <- factor(test$satisfaction,  levels = levels(train$satisfaction))
test$customer_type <- factor(test$customer_type, levels = levels(train$customer_type))
test$travel_type   <- factor(test$travel_type,   levels = levels(train$travel_type))
test$cabin_class   <- factor(test$cabin_class,   levels = levels(train$cabin_class))

n_train <- nrow(train)
n_test  <- nrow(test)
```

## Standardize test using train's center/scale

The held-out set's own mean/sd never touches model fitting or evaluation —
reusing train's center/scale is what makes this an honest leakage-free test.

```{r standardize}
standardize_with_train <- function(train_raw, test_raw) {
  ctr <- mean(train_raw); scl <- sd(train_raw)
  list(train_z = (train_raw - ctr) / scl, test_z = (test_raw - ctr) / scl)
}

fd  <- standardize_with_train(log1p(train$flight_distance), log1p(test$flight_distance))
dd  <- standardize_with_train(log1p(train$dep_delay),       log1p(test$dep_delay))
ade <- standardize_with_train(train$arr_delay_extra,        test$arr_delay_extra)
ag  <- standardize_with_train(train$age,                    test$age)
sc  <- standardize_with_train(train$seat_comfort,            test$seat_comfort)

train$flight_distance_z <- fd$train_z; test$flight_distance_z <- fd$test_z
train$dep_delay_z       <- dd$train_z; test$dep_delay_z       <- dd$test_z
train$arr_delay_extra_z <- ade$train_z; test$arr_delay_extra_z <- ade$test_z
train$age_z             <- ag$train_z; test$age_z             <- ag$test_z
train$seat_comfort_z    <- sc$train_z; test$seat_comfort_z    <- sc$test_z

extra_service_items <- c("time_convenient", "food_drink", "gate_location", "wifi",
                          "entertainment", "online_support", "ease_booking",
                          "onboard_service", "legroom", "baggage", "checkin",
                          "cleanliness", "online_boarding")

for (v in extra_service_items) {
  sv <- standardize_with_train(train[[v]], test[[v]])
  train[[paste0(v, "_z")]] <- sv$train_z
  test[[paste0(v, "_z")]]  <- sv$test_z
}
extra_service_z <- paste0(extra_service_items, "_z")
```

## Design matrices

Same formulas as Sections 2 and 3, applied to both train and test, with a
hard check that the resulting column names line up exactly before either
matrix touches a model.

```{r design-matrices}
flat_formula <- ~ age_z + seat_comfort_z + flight_distance_z + cabin_class +
                    dep_delay_z + arr_delay_extra_z + travel_type + customer_type

X_train_flat <- model.matrix(flat_formula, data = train)[, -1]
X_test_flat  <- model.matrix(flat_formula, data = test)[, -1]
stopifnot(identical(colnames(X_train_flat), colnames(X_test_flat)))

hs_formula <- as.formula(paste(
  "~ age_z + seat_comfort_z + flight_distance_z + cabin_class +",
  "dep_delay_z + arr_delay_extra_z + travel_type + customer_type +",
  paste(extra_service_z, collapse = " + ")
))

X_train_hs <- model.matrix(hs_formula, data = train)[, -1]
X_test_hs  <- model.matrix(hs_formula, data = test)[, -1]
stopifnot(identical(colnames(X_train_hs), colnames(X_test_hs)))

y_train <- as.numeric(train$satisfaction == "satisfied")
y_test  <- as.numeric(test$satisfaction  == "satisfied")
```

## Posterior draws: flat (fresh refit), horseshoe and BAS (reloaded)

The flat model is refit here without `log_lik` (not needed for prediction,
and it's the expensive part of the Section 5 version — dropping it keeps
this fast). The horseshoe posterior and the BAS fit are reused directly from
the `.rds` files Sections 3 and 4 already saved.

```{r fit-flat, cache=TRUE}
p_flat <- ncol(X_train_flat)

model_string_flat <- "
model {
  for (i in 1:ntr) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- beta0 + inprod(X[i,], beta[])
  }
  beta0 ~ dnorm(0, 0.01)
  for (j in 1:p_flat) { beta[j] ~ dnorm(0, 0.01) }
}
"
jm_flat <- jags.model(textConnection(model_string_flat),
                       data = list(y = y_train, X = X_train_flat, ntr = n_train, p_flat = p_flat),
                       n.chains = 4, n.adapt = 1000)
update(jm_flat, n.iter = 2000)
samples_flat <- coda.samples(jm_flat, variable.names = c("beta0", "beta"), n.iter = 2000)
flat_mat <- as.matrix(samples_flat)

beta0_flat <- flat_mat[, "beta0"]
beta_flat  <- flat_mat[, paste0("beta[", seq_len(p_flat), "]")]
```

```{r load-hs-bas}
samples_hs <- readRDS("data/samples_hs.rds")
hs_mat     <- as.matrix(samples_hs)
p_hs       <- ncol(X_train_hs)

beta0_hs <- hs_mat[, "beta0"]
beta_hs  <- hs_mat[, paste0("beta[", seq_len(p_hs), "]")]

bas_glm_fit <- readRDS("data/bas_glm_fit.rds")
```

## Posterior-mean predicted probability, averaged over draws

Done in batches over posterior draws rather than all at once — at
`r format(n_test, big.mark = ",")` test rows, materializing a full
`n_test`$\times S$ matrix for the flat model's 8,000 draws would run into
several GB in one allocation. Batching keeps memory bounded regardless of
how many draws either chain has.

```{r predict-function}
posterior_mean_prob <- function(X_test, beta0_samples, beta_samples, batch_size = 200) {
  n_test <- nrow(X_test)
  S <- length(beta0_samples)
  prob_sum <- numeric(n_test)
  starts <- seq(1, S, by = batch_size)
  for (s0 in starts) {
    s1  <- min(s0 + batch_size - 1, S)
    eta <- outer(rep(1, n_test), beta0_samples[s0:s1]) +
             X_test %*% t(beta_samples[s0:s1, , drop = FALSE])
    prob_sum <- prob_sum + rowSums(plogis(eta))
  }
  prob_sum / S
}

p_flat_test <- posterior_mean_prob(X_test_flat, beta0_flat, beta_flat)
p_hs_test   <- posterior_mean_prob(X_test_hs,   beta0_hs,   beta_hs)

# predict.bas already does BMA-averaging internally — no manual draw loop needed here
bas_pred   <- predict(bas_glm_fit, newdata = test, type = "response", estimator = "BMA")
p_bas_test <- bas_pred$fit
```

## Evaluation: accuracy and confusion matrix at a 0.5 cutoff

```{r evaluate}
evaluate_predictions <- function(p_hat, y_true) {
  pred <- ifelse(p_hat >= 0.5, 1, 0)
  cm  <- table(Predicted = pred, Actual = y_true)
  acc <- mean(pred == y_true)
  list(confusion = cm, accuracy = acc)
}

res_flat <- evaluate_predictions(p_flat_test, y_test)
res_hs   <- evaluate_predictions(p_hs_test,   y_test)
res_bas  <- evaluate_predictions(p_bas_test,  y_test)

cat("Flat model:\n");      print(res_flat$confusion)
cat("\nHorseshoe model:\n"); print(res_hs$confusion)
cat("\nBAS bas.glm:\n");     print(res_bas$confusion)

summary_df <- data.frame(
  Model    = c("Flat (9 predictors)", "Horseshoe (22 predictors)", "BAS bas.glm"),
  Accuracy = c(res_flat$accuracy, res_hs$accuracy, res_bas$accuracy)
)
print(summary_df)
```

```{r interpretation-setup, include=FALSE}
acc_tab    <- summary_df[order(-summary_df$Accuracy), ]
best_model <- acc_tab$Model[1]
best_acc   <- acc_tab$Accuracy[1]
worst_acc  <- acc_tab$Accuracy[nrow(acc_tab)]
acc_spread <- 100 * (best_acc - worst_acc)
```

## Results

On the held-out test set of `r format(n_test, big.mark = ",")` customers —
untouched during fitting, standardization, or variable selection — accuracy
at a 0.5 cutoff is `r sprintf("%.1f%%", 100 * res_flat$accuracy)` for the
flat model, `r sprintf("%.1f%%", 100 * res_hs$accuracy)` for the horseshoe
model, and `r sprintf("%.1f%%", 100 * res_bas$accuracy)` for BAS's
`bas.glm`. The best-performing model on this held-out set is
**`r best_model`**, at `r sprintf("%.1f%%", 100 * best_acc)`; the full
spread across all three is `r sprintf("%.1f", acc_spread)` percentage
points.

This is worth reading against Section 5 rather than in isolation: WAIC and
PSIS-LOO there favored the horseshoe decisively *in-sample*, on the
training data it was fit to. Whether that advantage survives contact with
genuinely new customers is exactly what this section tests — and the
honest reading is whichever way the numbers land here, not an assumption
that the WAIC winner must also win the accuracy comparison. A model with
13 extra predictors has more room to overfit the training subsample even
under a shrinkage prior; if its test accuracy comes in close to or below
the flat model's despite the large WAIC gap, that gap says more about
in-sample fit than out-of-sample generalization, and the report should say
so plainly rather than picking the in-sample winner by default.

The confusion matrices above give the raw counts behind each accuracy
figure — worth a glance for whether errors lean toward false positives or
false negatives for any one model, even without computing precision/recall
formally, since the outcome here is close to balanced and a lopsided error
pattern would be the more interesting thing to flag qualitatively.

## ROC and AUC

Accuracy only tells us how often the model lands on the right side of the
0.5 cutoff, which is an arbitrary choice. It ignores how confident the model
was.

The ROC curve tries every possible cutoff instead, and AUC is the area under
it. AUC measures how well the model *ranks* customers, with no threshold
involved: 0.5 is random, 1 is perfect.

```{r roc-auc, fig.width = 6.5, fig.height = 6.5}
library(pROC)

roc_flat <- roc(y_test, p_flat_test, quiet = TRUE)
roc_hs   <- roc(y_test, p_hs_test,   quiet = TRUE)
roc_bas  <- roc(y_test, p_bas_test,  quiet = TRUE)

draw_roc <- function() {
  plot(roc_flat, col = "steelblue", lwd = 2,
       main = "ROC curves on the held-out test set")
  lines(roc_hs,  col = "firebrick", lwd = 2)
  lines(roc_bas, col = "darkgreen", lwd = 2, lty = 2)
  legend("bottomright", bty = "n", lwd = 2,
         col = c("steelblue", "firebrick", "darkgreen"),
         lty = c(1, 1, 2),
         legend = c(sprintf("Flat (AUC %.3f)",      auc(roc_flat)),
                    sprintf("Horseshoe (AUC %.3f)", auc(roc_hs)),
                    sprintf("BAS (AUC %.3f)",       auc(roc_bas))))
}

draw_roc()

# also save for the report
dir.create("figs", showWarnings = FALSE)
png("figs/roc_curves.png", width = 6.5, height = 6.5, units = "in", res = 150)
draw_roc()
invisible(dev.off())

summary_df$AUC <- c(as.numeric(auc(roc_flat)),
                    as.numeric(auc(roc_hs)),
                    as.numeric(auc(roc_bas)))
print(summary_df)
```

AUC and accuracy can disagree. A model can rank customers well but still get
many of them wrong at 0.5, if all its probabilities are shifted up or down.
The next section checks for exactly that.

## Calibration

Accuracy and AUC both only look at the *order* of the predictions. But we
want the probability itself to mean something: out of the customers the
model says are 70% likely to be satisfied, about 70% should actually be
satisfied. A model can rank well and still be badly calibrated, which for a
Bayesian model is a real problem, not a detail.

To check it we split the test customers into deciles by predicted
probability, and in each bin compare the average prediction against what
actually happened. If the model is calibrated the points sit on the
diagonal.

```{r calibration, fig.width = 7, fig.height = 5.5}
calibration_bins <- function(p_hat, y_true, label, n_bins = 10) {
  # Bin on quantiles rather than fixed-width, so each bin holds a comparable
  # number of customers even where predicted probabilities pile up.
  brks <- unique(quantile(p_hat, probs = seq(0, 1, length.out = n_bins + 1)))
  bin  <- cut(p_hat, breaks = brks, include.lowest = TRUE, labels = FALSE)
  agg  <- aggregate(cbind(predicted = p_hat, observed = y_true),
                    by = list(bin = bin), FUN = mean)
  agg$n     <- as.numeric(table(bin))
  agg$Model <- label
  agg
}

calib <- rbind(
  calibration_bins(p_flat_test, y_test, "Flat (9 predictors)"),
  calibration_bins(p_hs_test,   y_test, "Horseshoe (22 predictors)"),
  calibration_bins(p_bas_test,  y_test, "BAS bas.glm")
)

p_calib <- ggplot(calib, aes(x = predicted, y = observed, colour = Model)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_line() +
  geom_point(aes(size = n), alpha = 0.8) +
  scale_colour_manual(values = c("Flat (9 predictors)"       = "steelblue",
                                  "Horseshoe (22 predictors)" = "firebrick",
                                  "BAS bas.glm"               = "darkgreen")) +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(title = "Calibration on the held-out test set",
       subtitle = "Deciles of predicted probability; points on the diagonal are perfectly calibrated",
       x = "Mean predicted probability (within bin)",
       y = "Observed satisfaction rate (within bin)",
       size = "customers in bin") +
  theme_minimal()

ggsave("figs/calibration.png", p_calib, width = 7, height = 5.5, dpi = 150)
p_calib
```

The same thing as a single number: the average gap between predicted and
observed rate across the bins, weighted by bin size. This is usually called
the expected calibration error (ECE). Lower is better, 0 is perfect.

```{r calibration-error}
ece <- sapply(split(calib, calib$Model), function(d)
  sum(d$n * abs(d$predicted - d$observed)) / sum(d$n))

round(data.frame(ECE = sort(ece)), 4)
```

Two things worth checking here. If a model's curve sits consistently below
the diagonal it is overstating satisfaction, and above means it understates
it. And it is worth seeing whether the model that wins on accuracy also wins
on calibration, since they measure different things: ranking customers needs
good AUC, but saying "this customer has a 72% chance" needs good calibration.
