---
title: "00 - Data Audit, EDA, and the Subsampling Decision 01"
subtitle: "Airline Customer Satisfaction -- Bayesian Learning & Monte Carlo Simulations, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                       fig.align = "center", cache = FALSE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(caret)
library(RColorBrewer)
```

# Data audit, EDA, and the subsampling decision

## Loading the data and basic recoding

We work with the full \~130k-row source file rather than the curated
1,000-row `airline_sub.csv` described in the brief (see the sampling
decision below for how we reconcile the two).

```{r load-data}
raw <- read.csv("../data/Airline_customer_satisfaction.csv", stringsAsFactors = FALSE)

df <- raw %>%
  mutate(
    satisfaction   = factor(satisfaction, levels = c("dissatisfied", "satisfied")),
    Class          = factor(Class, levels = c("Eco", "Eco Plus", "Business"), ordered = TRUE),
    Customer.Type  = factor(Customer.Type),
    Type.of.Travel = factor(Type.of.Travel)
  )

n_total   <- nrow(df)
sat_rate  <- mean(df$satisfaction == "satisfied")
```

The dataset contains `r format(n_total, big.mark = ",")` customers. The
overall satisfaction rate is `r sprintf("%.1f%%", 100 * sat_rate)`,
balanced enough for a straightforward Bernoulli / logistic treatment
later on. `Class` has 3 levels (Eco, Eco Plus, Business);
`Customer.Type` and `Type.of.Travel` are binary.

## Correlation structure of the 14 service-quality items

The Likert-style service ratings (0-5) are not independent. We check
this with a correlation heatmap, ordered by hierarchical clustering so
that correlated blocks are visually adjacent.

```{r service-items}
service_items <- c("Seat.comfort", "Departure.Arrival.time.convenient", "Food.and.drink",
                    "Gate.location", "Inflight.wifi.service", "Inflight.entertainment",
                    "Online.support", "Ease.of.Online.booking", "On.board.service",
                    "Leg.room.service", "Baggage.handling", "Checkin.service",
                    "Cleanliness", "Online.boarding")

cor_services <- cor(df[, service_items], use = "pairwise.complete.obs")
```

```{r corr-heatmap, fig.width=8.5, fig.height=7.5}
corrplot(cor_services, method = "color", type = "upper", order = "hclust",
         col = colorRampPalette(rev(brewer.pal(9, "RdBu")))(200),
         tl.col = "black", tl.srt = 45, tl.cex = 0.75,
         addCoef.col = "black", number.cex = 0.55,
         diag = FALSE,
         title = "Correlation matrix of the 14 service-quality items\n(hierarchical clustering order)",
         mar = c(0, 0, 2, 0))
```

```{r cluster-ranges}
cluster_range <- function(items) {
  sub <- cor_services[items, items]
  vals <- sub[upper.tri(sub)]
  sprintf("%.2f-%.2f", min(vals), max(vals))
}

cluster_booking  <- c("Inflight.wifi.service", "Online.support",
                       "Ease.of.Online.booking", "Online.boarding")
cluster_onboard  <- c("On.board.service", "Leg.room.service",
                       "Baggage.handling", "Cleanliness")
cluster_catering <- c("Seat.comfort", "Food.and.drink",
                       "Departure.Arrival.time.convenient", "Gate.location")

r_booking  <- cluster_range(cluster_booking)
r_onboard  <- cluster_range(cluster_onboard)
r_catering <- cluster_range(cluster_catering)
```

Three blocks emerge clearly:

- **Booking / connectivity**: `Inflight wifi service`, `Online support`,
  `Ease of Online booking`, `Online boarding` move together (pairwise
  correlations `r r_booking`).
- **On-board comfort**: `On-board service`, `Leg room service`,
  `Baggage handling`, `Cleanliness` form a second cluster
  (`r r_onboard`).
- **Seat / catering / ground experience**: `Seat comfort`,
  `Food and drink`, `Departure/Arrival time convenient`, `Gate location`
  form a third, slightly tighter cluster (`r r_catering`) -- this one
  was not obvious from a first pass and only shows up once the full
  correlation matrix is inspected; it is also, as the next section
  shows, exactly the cluster that carries most of the structural-zero
  ambiguity.

This is the empirical basis for using a shrinkage prior (Section 3)
rather than treating all 14 items as freely and independently
informative.

## Structural zeros in the service ratings

Several items show a suspiciously large share of exact `0` ratings,
which plausibly encode "not applicable" (e.g. no on-board wifi on some
flights) rather than a genuine bottom-of-scale rating.

```{r structural-zeros}
zero_share <- sort(sapply(df[, service_items], function(x) mean(x == 0)),
                    decreasing = TRUE)

knitr::kable(
  data.frame(item = names(zero_share), zero_share = round(zero_share, 3)),
  row.names = FALSE, col.names = c("Service item", "Share of rows = 0"),
  caption = "Share of rows recording a 0 rating, by service item"
)
```

We flag, but do not model, the items with a non-trivial zero share
(above 2%) as a single per-row indicator, to be used later only as a
robustness check (Section 5: does dropping these rows change the
conclusions?).

```{r flag-structural-zero}
zero_candidates <- names(zero_share)[zero_share > 0.02]

df <- df %>%
  mutate(structural_zero_flag = if_any(all_of(zero_candidates), ~ .x == 0))

n_flagged <- sum(df$structural_zero_flag)
```

`r length(zero_candidates)` items are flagged
(`r paste(zero_candidates, collapse = ", ")`), affecting
`r format(n_flagged, big.mark = ",")` rows
(`r sprintf("%.1f%%", 100 * mean(df$structural_zero_flag))` of the
data). This matches the earlier finding that these are exactly the items
in the "seat / catering / ground experience" cluster above -- consistent
with a shared "not applicable on this flight" mechanism rather than four
unrelated data-quality issues.

## Decorrelating Departure and Arrival delay

```{r delay-collinearity}
delay_cor <- cor(df$Departure.Delay.in.Minutes, df$Arrival.Delay.in.Minutes,
                  use = "complete.obs")
n_missing_arr <- sum(is.na(df$Arrival.Delay.in.Minutes))
```

`Departure Delay` and `Arrival Delay` correlate at
`r sprintf("%.3f", delay_cor)`: using both raw would be close to
duplicating one variable. `Arrival Delay` also has `r n_missing_arr`
missing values (`r sprintf("%.2f%%", 100 * n_missing_arr / n_total)`).

Rather than a simple difference, we regress `Arrival Delay` on
`Departure Delay` and keep the **residual** as a second covariate -- by
construction uncorrelated with `Departure Delay`, and interpretable as
"minutes gained or lost in the air relative to what the departure delay
alone would predict":

```{r delay-residual}
delay_fit <- lm(Arrival.Delay.in.Minutes ~ Departure.Delay.in.Minutes, data = df)

df$ArrivalDelay.Residual <- NA_real_
df$ArrivalDelay.Residual[!is.na(df$Arrival.Delay.in.Minutes)] <- residuals(delay_fit)

check_cor <- cor(df$Departure.Delay.in.Minutes, df$ArrivalDelay.Residual,
                  use = "complete.obs")
```

The fitted relationship is Arrival $\approx$
`r sprintf("%.2f", coef(delay_fit)[1])` $+$
`r sprintf("%.3f", coef(delay_fit)[2])` $\times$ Departure (R$^2$ =
`r sprintf("%.3f", summary(delay_fit)$r.squared)`), and the residual is
essentially uncorrelated with Departure Delay by construction (r =
`r formatC(check_cor, format = "e", digits = 2)`). The `r n_missing_arr`
missing arrival delays propagate to missing residuals here; we do not
drop or impute them in this section -- Section 2 treats them properly as
stochastic nodes inside the JAGS model rather than discarding rows.

## Stratified subsample: matching the brief's intended scale

The brief describes a curated 1,000-row file; we have the full
`r format(n_total, big.mark = ",")`-row source. We draw a stratified
training subsample of $n \approx 1{,}200$ (stratified jointly on
`satisfaction` and `Class`, matching the brief's intended scale while
keeping MCMC fast in later sections) and keep essentially all remaining
rows as a held-out test set for prediction (Section 6).

```{r stratified-sample}
set.seed(123)

strata <- interaction(df$satisfaction, df$Class, drop = TRUE)
train_idx <- createDataPartition(y = strata, p = 1200 / n_total, list = FALSE)

train_df <- df[train_idx, ]
test_df  <- df[-train_idx, ]
```

Training subsample: `r nrow(train_df)` rows. Held-out test set:
`r format(nrow(test_df), big.mark = ",")` rows.

```{r stratification-check}
compare_tbl <- function(var) {
  full  <- prop.table(table(df[[var]]))
  train <- prop.table(table(train_df[[var]]))
  round(cbind(full = full, train = train), 3)
}

cat("Satisfaction proportions (full data vs. training subsample):\n")
print(compare_tbl("satisfaction"))

cat("\nClass proportions (full data vs. training subsample):\n")
print(compare_tbl("Class"))

cat("\nCustomer Type proportions (full data vs. training subsample):\n")
print(compare_tbl("Customer.Type"))
```

The stratified draw preserves the population proportions closely on all
three grouping variables, so subsequent sections can treat `train_df` as
a faithful, fast-to-fit stand-in for the full population for model
development, with `test_df` reserved untouched for genuine out-of-sample
evaluation.

## Service ratings by satisfaction

```{r boxplots-by-satisfaction, fig.width=9, fig.height=7}
df %>%
  select(satisfaction, all_of(service_items)) %>%
  pivot_longer(-satisfaction, names_to = "item", values_to = "rating") %>%
  ggplot(aes(x = satisfaction, y = rating, fill = satisfaction)) +
  geom_boxplot(outlier.size = 0.4, alpha = 0.85) +
  facet_wrap(~ item, ncol = 4, scales = "free_y") +
  scale_fill_manual(values = c("dissatisfied" = "#D85A30", "satisfied" = "#3B6D11")) +
  labs(title = "Service-quality ratings by satisfaction outcome",
       x = NULL, y = "Rating (0-5)") +
  theme_minimal(base_size = 9) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 0),
        strip.text = element_text(face = "bold", size = 7.5))
```

Satisfied customers rate essentially every service item higher, most
visibly for `Online boarding`, `Inflight entertainment` and
`Ease of Online booking` -- consistent with the booking/connectivity
cluster identified above being a strong candidate driver, to be
confirmed formally in the modelling sections.

## Saving the derived datasets for the rest of the analysis

```{r save-outputs}
saveRDS(df, "airline_full_clean.rds")
write.csv(train_df, "airline_train_n1200.csv", row.names = FALSE)
write.csv(test_df,  "airline_test_holdout.csv", row.names = FALSE)
```

`airline_full_clean.rds` (full data, recoded factors,
`ArrivalDelay.Residual` and `structural_zero_flag` added),
`airline_train_n1200.csv` and `airline_test_holdout.csv` are written
next to this file and are the inputs for every section from here on --
no section past this point should re-derive these variables from the raw
file. **Keep this Rmd and all the `.csv`/`.rds` files in the same
folder** (no subfolders) so the relative paths resolve. Section 1 onward
each live in their own numbered `.Rmd` file in this same folder and read
these outputs back in rather than re-deriving them.
